---
title: "Syllabus Electiva II: Negociación de Conflictos Module A AE"
author: "Distance Learning Faculty (FAEDIS)"
date: "2022-1"
output:
  pdf_document:
    number_sections: yes
  html_document:
    df_print: paged
    highlight: tango
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 4
bibliography: conflict_neg.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{r}

library(tidyverse)
library(knitr)
```

# Professor and meeting hours with students

- __Name__

    + Luis Francisco Gomez Lopez

- __Communication through the virtual classroom__

    + Every Sunday between 07:00 hours and 08:00 hours: 
    
        + _Primer Corte 30% > Virtual Meetings > Meeting hours with students_
    
    + Doubts and concerns of the topics of the course, please use the _Forum: Questions or concerns_ that can be accessed at the top of the platform
    
        + In this way, doubts can be shared with all students and all course participants can contribute to the learning process
    
    + Complaints regarding grades, please send an email to __luis.gomezl@unimilitar.edu.co__
    
        + To make any complaints, you are asked to wait first for the grades to be published on the News Bulletin Board and to be made based on the feedbacks that will be made taking into account the evaluation parameters
        
        + Urgent issues, you can call me to +57 3124541536 Monday, Wednesday, and Friday between 14:00 hours and 15:00 hours. Please do not send me whatsapp messages or call me on weekends.

# Presentation and objectives of the subject

## Description

- Credits: 4
- Type: Theoretical-Practice
- Area of Knowledge: Professional
- Semester: 8
- Block: 2 (March 28 – June 03, 2021) 

## Justification

Conflicts are frequent situations in the management of organizations. Due to the above, business administrators require guidance to understand business relationships in order to persuade other parties under ethical standards through negotiation tactics that mitigate the negative effects of the conflict.
To achieve this, it is necessary to develop in the future business administrator skills and abilities to lead groups and organizations using tools that allow him to communicate effectively and manage conflicts strategically between internal and external members of an organization. 

In that sense, it is sought that the future business administrator consolidate the necessary skills to be a better negotiator and reach beneficial agreements for the parties involved in negotiation and conflict situations. 

## General objective

Develop in students, negotiation conflict skills in diverse scenarios through the review, analysis and application of negotiation methods and techniques.

# Competencies

- Global competency:

    + Resolve conflict situations strategically and ethically in the organization and its environment to generate social, economic or environmental benefits.

- Specific competencies:

    + Lead the human talent solving conflicts in a strategic way to achieve the common objectives of the organization. 
    
    + Recognize the importance of ethics in the negotiation of conflicts to generate long-term relationships with the stakeholders of the organization. 
    
    + Plan negotiation strategies based on information from the environment to obtain collective benefits.

- Competencies SABE curriculum

    + Skills in handling a second language 
    + Ability to work in teams
    + Dialogic and deliberative capabilities
    + Use of argumentative thinking
    
# Schedule and contents

## Reading schedule

```{r}

lewicki <- rep("[@lewicki_essentials_2016]", times = 3)

tibble(Week = c("1","2","3", "3", "4", "5", "6", "6", "7", "8", "9", "9"),
       Material        = rep(c(lewicki, 
                               "[@program_on_negotiation_three_2008]"), 
                             times = 3),
       `Chapter/Round` = c("1","2","3", "I", "4", "6", "7", "II", "8", "5", "12", "III"),
       `Initial page`  = c("1", "28", "60", " ", "89", "139", "165", " ", "182", "114", "273", " "),
       `Final page`    = c("27", "59", "88", " ", "113", "164", "181", " ", "202", "138", "282", " "),
       `Number of pages` = c("27", "32", "29", " ", "25", "26", "17", " ", "21", "25", "10", " ")) %>% 
    kable()
```

## Detailed schedule of virtual meetings and activities

### Primer corte 30%

- March 28

    + __Virtual meeting Presentation Syllabus__
        + __Meeting time__: 19:00 - 19:45 hours
        + __Prerequisites__:
            + Review the course program in advance, which can be downloaded by following the route _Primer corte 30% > Support material - Primer corte 30% > 000_syllabus_conflict_neg_2021_1
            + Have a headset with a built-in microphone to isolate noise from the environment
            
- March 30

    + __Virtual meeting The Nature of Negotiation__
        + __Meeting time__: 19:00 - 19:45 hours 
        + __Prerequisites__: 
            + Read and review [@lewicki_essentials_2016, Chapter 1]
            + Have a headset with a built-in microphone to isolate noise from the environment

- April 4

    + __Questionnaire 1 Opening__
        + __Opening time__: 00:00 hours
        + __Prerequisites__:
            + Read and review [@lewicki_essentials_2016, Chapters 1 & 2]

    + __Virtual meeting Strategy and Tactics of Distributive Bargaining__
        + __Meeting time__: 19:00 - 19:45 hours
        + __Prerequisites__:  
            + Read and review [@lewicki_essentials_2016, Chapter 2]
            + Have a headset with a built-in microphone to isolate noise from the environment
            
- April 10

    + __Questionnaire 1 Closing__
        + __Closing time__: 22:00 hours
        + __Prerequisites__: 
            + Have answered the Questionnaire 1 before 22:00 hours
            
- April 11 - 17 

    + Holy Week
            
- April 18 

    + __Instructions Opening Midterm Exam Three Party Coalition Exercise Round 1__
        + __Opening time__: 00:00 hours
        + __Prerequisites__:
            + Read and review role play instructions [@program_on_negotiation_three_2008]

    + __Virtual meeting Strategy and Tactics of Integrative Negotiation and Three-Party Coalition Exercise Round 1__
        + __Meeting time__: 19:00 - 19:45 hours
        + __Prerequisites__: 
            + Read and review role play instructions [@program_on_negotiation_three_2008]
            + Read and review [@lewicki_essentials_2016, Chapter 3]
            + Have a headset with a built-in microphone to isolate noise from the environment
            
- April 23

    + __Midterm Exam Three Party Coalition Exercise Round 1__
        + __Opening time__: 15:30 hours
        + __Closing time__: 17:30 hours
        + __Prerequisites__: 
            + Read and review role play instructions [@program_on_negotiation_three_2008]
            + Read and review [@lewicki_essentials_2016, Chapter 3]
            + Have a headset with a built-in microphone to isolate noise from the environment

### Segundo corte 30%

- April 25

    + __Questionnaire 2 Opening__
        + __Opening time__: 00:00 hours
        + __Prerequisites__:
            + Read and review [@lewicki_essentials_2016, Chapters 4 & 6]

    + __Virtual meeting Negotiation: Strategy and Planning__
        + __Meeting time__: 19:00 - 19:45 hours
        + __Prerequisites__:  
            + Read and review [@lewicki_essentials_2016, Chapter 4]
            + Have a headset with a built-in microphone to isolate noise from the environment
            
- May 2

    + __Virtual meeting Perception, Cognition, and Emotion__
        + __Meeting time__: 19:00 - 19:45 hours
        + __Prerequisites__:  
            + Read and review [@lewicki_essentials_2016, Chapter 6]
            + Have a headset with a built-in microphone to isolate noise from the environment
            
- May 8

    + __Questionnaire 2 Closing__
        + __Closing time__: 22:00 hours
        + __Prerequisites__: 
            + Have answered the Questionnaire 2 before 22:00 hours
            
- May 9 

    + __Instructions Opening Midterm Exam Three Party Coalition Exercise Round 2__
        + __Opening time__: 00:00 hours
        + __Prerequisites__:
            + Read and review role play instructions [@program_on_negotiation_three_2008]

    + __Virtual meeting Communication and Three-Party Coalition Exercise Round 2__
        + __Meeting time__: 19:00 - 19:45 hours
        + __Prerequisites__: 
            + Read and review role play instructions [@program_on_negotiation_three_2008]
            + Read and review [@lewicki_essentials_2016, Chapter 7]
            + Have a headset with a built-in microphone to isolate noise from the environment
            
- May 14

    + __Midterm Exam Three Party Coalition Exercise Round 2__
        + __Opening time__: 15:30 hours
        + __Closing time__: 17:30 hours
        + __Prerequisites__: 
            + Read and review role play instructions [@program_on_negotiation_three_2008]
            + Read and review [@lewicki_essentials_2016, Chapter 7]
            + Have a headset with a built-in microphone to isolate noise from the environment

### Tercer corte 40%

- May 16

    + __Questionnaire 3 Opening__
        + __Opening time__: 00:00 hours
        + __Prerequisites__:
            + Read and review [@lewicki_essentials_2016, Chapters 8 & 5]

    + __Virtual meeting Negotiation: Finding and Using Negotiation Power__
        + __Meeting time__: 19:00 - 19:45 hours
        + __Prerequisites__:  
            + Read and review [@lewicki_essentials_2016, Chapter 8]
            + Have a headset with a built-in microphone to isolate noise from the environment
            
- May 18

    + __Virtual meeting Ethics in Negotiation__
        + __Meeting time__: 19:00 - 19:45 hours
        + __Prerequisites__:  
            + Read and review [@lewicki_essentials_2016, Chapter 5]
            + Have a headset with a built-in microphone to isolate noise from the environment
            
- May 22

    + __Questionnaire 3 Closing__
        + __Closing time__: 22:00 hours
        + __Prerequisites__: 
            + Have answered the Questionnaire 2 before 22:00 hours

- May 23

    + __Instructions Opening Final Exam Three Party Coalition Exercise Modified__
        + __Opening time__: 00:00 hours
        + __Prerequisites__:
            + Read and review role play instructions [@program_on_negotiation_three_2008]

    + __Virtual meeting Best Practices in Negotiations and Three-Party Coalition Exercise Modified__
        + __Meeting time__: 19:00 - 19:45 hours
        + __Prerequisites__: 
            + Read and review role play instructions [@program_on_negotiation_three_2008]
            + Read and review [@lewicki_essentials_2016, Chapter 12]
            + Have a headset with a built-in microphone to isolate noise from the environment
            
- May 28 

    + __Final Exam Three Party Coalition Exercise Modified__
        + __Opening time__: 15:30 hours
        + __Closing time__: 17:30 hours
        + __Prerequisites__: 
            + Read and review role play instructions [@program_on_negotiation_three_2008]
            + Read and review [@lewicki_essentials_2016, Chapter 12]
            + Have a headset with a built-in microphone to isolate noise from the environment
            
# Qualifiable activities, weights and process in case of no presentation

## Activities and weights

- __Primer corte__: 30%

    + __Questionnaire 1__: 15% 
    + __Midterm Exam Three Party Coalition Exercise Round 1__: 15% 

- __Segundo corte__: 30%

    + __Questionnaire 2__: 15% 
    + __Midterm Exam Three Party Coalition Exercise Round 2__: 15%

- __Tercer corte__: 40%

    + __Questionnaire 3__: 15%
    + __Final Exam Three Party Coalition Exercise Modified__: 25%
    
## Process in case of no presentation

- Students who do not present a __Questionnaire__ due to force majeure or illness, must send the respective excuse scanned to __luis.gomezl@unimilitar.edu.co__. If it is not sent, the note will be zero. Once the excuse has been satisfactorily received, the student should __not__ present the activity and the corresponding grade will be equal to the immediately subsequent grade obtained in the __Midterm Exam 1__ or __Midterm Exam 2__ or __Final Exam__ according to the case.

- Students who do not present a __Midterm Exam 1__ or __Midterm Exam 2__ or the __Final Exam__, must send and email to __admon.distancia@unimilitar.edu.co__ two days before or two days after the closing date of these activities. However, this request can be made only once during the course. If it is not requested under these conditions, the grade will be zero.

# Study material

- [@lewicki_essentials_2016, Chapters 1, 2, 3, 4, 6, 7, 8, 5 & 12]

    + The book can be consulted through the virtual classroom
    
- [@program_on_negotiation_three_2008]

    + The document can be consulted through the virtual classroom
    
# About plagiarism

- __Acuerdo 02 de 2015 (7 de Mayo): Artículo 68. Escala de Calificaciones Numeral f__ 

_"La comprobación de todo fraude o intento de fraude en cualquier tipo de prueba, acarreará al estudiante y a su(s) colaborador(es), una calificación de CERO PUNTO CERO (0.0) en la prueba respectiva. Como quiera que el fraude o intento de cometerlo es una falta grave, el docente debe informar de este hecho al Director del programa respectivo, quien enviará el informe a la Decanatura de la Facultad, con el fin de iniciar los trámites correspondientes para aplicar la sanción disciplinaria prevista en éste reglamento.”_

- __Acuerdo 05 de 2018 (6 de Junio): Artículo 20 que modifica el Artículo 107 Faltas Disciplinarias Numeral c. Número 6 del Acuerdo 02 de 2015 (7 de Mayo)__

_"Hacer plagio en cualquier tipo de documento o desplegar alguna conducta dolosa que viole el régimen de propiedad intelectual de la Universidad."_

# Study material references
