---
title: "The Nature of Negotiation"
author: "Luis Francisco Gomez Lopez"
date: 2021-04-05 10:17:36 GMT -05:00
output:
  beamer_presentation:
    colortheme: dolphin
    fonttheme: structurebold
    theme: AnnArbor
    fig_caption: true
  slidy_presentation: default
bibliography: conflict_neg.bib
link-citations: yes
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{fontawesome5}
  - \usepackage{xcolor}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      message   = FALSE,
                      fig.align = "center")
```

```{r libraries}

library(tidyverse)
library(tidyquant)
```

# Contents

- Please Read Me
- Purpose
- Definition of negotiation
- Elements of negotiation situations
- Interdependence
- Approaches and tactics to negotiation
- Conflict
- Acknowledgments
- References

# Please Read Me

- Check the message __Welcome greeting__ published in the News Bulletin Board.

- Dear student please edit your profile uploading a photo where your face is clearly visible.

- The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

- This presentation is based on [@lewicki_essentials_2016, Chapter 1]

# Purpose

Understand the definition of negotiation, the elements in general of this process and the main types of negotiation that have been identified in the literature as well as the relationship between negotiation and conflict management.

# Definition of negotiation

- Definition of negotiation that is going to be adopted in the course:

    + " *a form of __decision making__ in which two or more parties talk with one another in an effort to __resolve their opposing interests__.*" [@lewicki_essentials_2016, Chapter 1, p 3]
    
    + Also check out in the __Links of interest__ the video
    
        + _How to negotiate properly?_[^1] [@magic_markers_como_2018]
    
[^1]: The video is in spanish

# Elements of negotiation situations

- 6 characteristics are mentioned in [@lewicki_essentials_2016, Chapter 1, p 6-9] 

    + Element 1 assumes that in a negotiation there are two or more parties involved (people, groups or organizations).
    
    + Elements 2, 3, 4 and 5 implies that there exist __interdependence__ between the parties involved[^2]: 
    
        + Parties depend on each other to achieve their own preferred outcome [@lewicki_essentials_2016, Chapter 1, p 10]
        
    + Element 6 considers that there are tangible and intangible aspects:
    
        + __Tangibles__: aspects of which it is sought to reach an agreement within the negotiation (prices, terms of a contract, product specifications).
        
        + __Intangibles__: psychological motivations implicit in a negotiation (personal values and emotions of the parties involved).
        
[^2]: Dear student if you don’t need to jointly agree with other parties to achieve your goals please don’t negotiate with them. Use another form of __decision making process__!

# Interdependence

- The type of interdependence that occurs between the parties affects the dynamics and the results of a negotiation:

    + __Zero-sum or distributive situation__: for one party to obtain a profit,it is necessary for another party to obtain a loss.
    
    + __Non-zero-sum or integrative situation__: there exists the __*possibility*__ that all parties involved can obtain a profit without a party obtaining a loss.

- The alternatives shape the interdependence where one way to analyze the alternatives is through the concept of:

    + __Best Alternative to a Negotiated Agreement (BATAN)__[^3]
    
[^3]: For more details check out [@program_of_negotiation_batna_2012]

# Interdependence

- Because the parties are interdependent during a negotiation there is a process called __mutual adjustment__ where the parties try to influence one another to reach an acceptable agreement.

    + At the beginning each party proposes an initial level on the tangible aspects of negotiation.
    
    + Then there are replicas and concessions where the __bargaining range__ tends to be reduced.
    
    + If the __bargaining range__ is reduced enough to become a point, an agreement is reached.

# Approaches and tactics to negotiation

- __Distributive Negotiation Approach__: it is conceived that there is only one winner in a particular situation and a course of action is sought to be the winner.

    + Tactic: __claim value__[^4] by doing whatever is necessary to claim the reward or gain the largest piece possible [@lewicki_essentials_2016, Chapter 1, p 16]
    
+ __Integrative Negotiation Approach__: attempts are made to find solutions so that both parties feel satisfied and reach their goals.
    
    + Tactic: __create value__[^5] by finding the way for the parties involved to achieve their objectives by obtaining more resources
    
[^4]: For more details check out [@spangler_creating_2003]
[^5]: For more details check out [@spangler_creating_2003]

# Conflict

- For definitions of __conflict__ please check out [@lewicki_essentials_2016, Chapter 1, p 18] 

    + 3 different definitions that are related are mentioned
    
- For the different levels of __conflict__ please check out [@lewicki_essentials_2016, Chapter 1, p 19] 

    + 4 levels on conflict are mentioned
    + We will not cover the first one in the course: 
    
      + _Intrapersonal or intrapsychic conflict_
      
- For dysfunctions and functions of __conflict__ please check out [@lewicki_essentials_2016, Chapter 1, p 19-21] 

    + Dysfunctions: negative aspects of conflict
    + Functions: productive aspects of conflict
    
# Conflict

```{r, out.width="90%"}

# Data
conflic_diag <- tibble(Dimension   = rep(c("Simmetry of conflict",
                           "Involment of third parties",
                           "Structure of the parties",
                           "Interaction",
                           "Interdependence",
                           "Size of stakes",
                           "Issue in question"), times = 2),
       Level       = c(rep("Easy to resolve", times = 7),
                       rep("Difficult to Resolve", times = 7)),
       label_text  = c("Equal damage",
                       "Trusted third parties",
                       "Strong Leadership",
                       "Long Term Relationship",
                       "Integrative situation",
                       "Small consequences",
                       "Divisible Issue",
                       "Asimmetry in damage suffered",
                       "No third parties",
                       "Disorganized",
                       "Single interaction",
                       "Distributive situation",
                       "Big consequences",
                       "Fundamental principles")) %>% 
    mutate(Level = as.factor(Level))

# Data auxiliary
conflic_diag_level <- conflic_diag %>% 
    slice(1:7) %>%
    select(-label_text) %>% 
    set_names(nm = c("Dimension", "Level_easy")) %>% 
    left_join(y = conflic_diag %>% 
                  slice(8:14) %>%
                  select(-label_text) %>% 
                  set_names(nm = c("Dimension", "Level_difficult")),
                            by = "Dimension")
    
conflic_diag %>% 
    ggplot() +
    geom_segment(data = conflic_diag_level,
                 aes(x = Level_easy, xend = Level_difficult,
                     y = Dimension,  yend = Dimension)) +
    geom_label(aes(x = Level, y = Dimension, label = label_text, fill = Dimension), 
               show.legend = FALSE,
               fontface    = "bold",
               size        = 3, 
               nudge_y     = 0.4,
               alpha       = 0.4) +
    geom_point(aes(x = Level, y = Dimension, color = Level),
               size  = 5) + 
    scale_color_tq() +
    scale_fill_manual(values = palette_light()[4:10] %>% as.vector()) +
    labs(x       = "",
         title   = "Conflict Diagnostic Model",
         caption = "Source: (Lewicki, Barry, and Saunders 2016, Chapter 1, p 22)") +
    theme(panel.border      = element_rect(fill = NA, color = "black"),
          plot.background   = element_rect(fill = "#f3fcfc"),
          panel.background  = element_rect(fill = "#f3f7fc"),
          legend.background = element_rect(fill = "#f3fcfc"),  
          plot.title        = element_text(face = "bold"), 
          axis.text.x       = element_blank(), 
          axis.ticks.x      = element_blank(),
          axis.title.y      = element_text(face = "bold"),
          legend.title      = element_text(face = "bold"),
          legend.position   = "bottom", 
          axis.text         = element_text(face = "bold"))
```

# Acknowledgments

- To my family that supports me

- To the taxpayers of Colombia and the __[UMNG students](https://www.umng.edu.co/estudiante)__ who pay my salary

- To the __[Business Science](https://www.business-science.io/)__ and __[R4DS Online Learning](https://www.rfordatasci.com/)__ communities where I learn __[R](https://www.r-project.org/about.html)__ 

- To the __[R Core Team](https://www.r-project.org/contributors.html)__, the creators of __[RStudio IDE](https://rstudio.com/products/rstudio/)__ and the authors and maintainers of the packages __[tidyverse](https://CRAN.R-project.org/package=tidyverse)__, __[tidyquant](https://CRAN.R-project.org/package=tidyquant)__ and __[tinytex](https://CRAN.R-project.org/package=tinytex)__ for allowing me to access these tools without paying for a license

- To the __[Linux kernel community](https://www.kernel.org/category/about.html)__ for allowing me the possibility to use some __[Linux distributions](https://static.lwn.net/Distributions/)__ as my main __[OS](https://en.wikipedia.org/wiki/Operating_system)__ without paying for a license

# References