---
title: The Nature of Negotiation
subtitle: Session 1
author:
  - name: 
      given: Luis F.
      family: Gómez L.
    orcid: 0000-0002-2877-9015
    email: luis.gomezl@unimilitar.edu.co
    affiliations:
      - name: Distance Learning Faculty
toc: true
toc-depth: 1
slide-level: 2
format:
  presentation-umng-revealjs:
    logo-vertical-strip: 000_logos/logo-white-vertical.png 
    embed-resources: true
    mermaid-format: png
    reference-location: document
date: last-modified
date-format: "D MMMM, YYYY"
knitr: 
  opts_chunk: 
    echo: false
    warning: false
    fig-align: center
    out-width: 85%
lang: en
bibliography: conflict_neg.bib
csl: apa-7.csl
editor: source
---

```{r}
#| label: libraries

library(tidyverse)
library(ggnewscale)
```

```{r}
#| label: palette

umng_palette <- c(
  "#043074",
  "#fdc600",
  "#0d5e30",
  "#ee2a24",
  "#fc6700",
  "#00b3f0",
  "#6e3c1a",
  "#f8941c",
  "#8e44ad",
  "#2c3e50",
  "#16a085",
  "#c0392b",
  "#e91e63"
)
```

# Please Read Me

##

-   Check the message **Welcome greeting** published in the News Bulletin Board.

-   Dear student please edit your profile uploading a photo where your face is clearly visible.

-   The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

-   This presentation is based on [@lewicki_negociacion_2024, Chapter 1]

## 

- **Purpose**

    - Understand the definition of negotiation, the elements in general of this process and the main types of negotiation that have been identified in the literature as well as the relationship between negotiation and conflict management.

# The roles managers play, negotiation and conflict resolution

##

```{mermaid}
%%| label: fig-manager-roles
%%| fig-cap: The roles managers play [@openstax_featured_principles_2019, chap. 2, sec. 1.2]

flowchart TB
    A("`Manager
    roles`") --> A1(Informational)
    A1 --> A11("`Monitor
    Disseminator
    Spokesperson`")
    A --> A2(Interpersonal)
    A2 --> A21("`Figurehead
    Leader
    Liason`")
    A --> A3(Decisional)
    A3 --> A31("`Entrepreneur
    **Disturbance handler**
    Resource allocator
    **Negotiator**`")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A2,A21,A3,A31 UmngNode;
```

##

- Management roles that apply directly to negotiation and conflict resolution (See [@mintzberg_mintzberg_1989, chap. 1])

    - **Disturbance handler**
    
        - Deals with conflicts or crises within the organization 
    
    - **Negotiator**
    
        - Engages in negotiations both within and outside the organization

# Elements of negotiation

##

- Definition of negotiation

    -   *A form of **decision making** in which 2 or more parties talk with one another in an effort to **resolve their opposing interests**.* [@lewicki_negociacion_2024, Chapter 1, p. 3]

    -   Also check out in the **Links of interest** the video

        -   *How to negotiate properly?*[^1] [@magic_markers_como_2018]

[^1]: The video is in spanish

## {.smaller}

-   6 elements of negotiation situations are mentioned in [@lewicki_negociacion_2024, Chapter 1, pp. 6-9]

    -   Element 1 assumes that in a negotiation there are two or more parties involved (people, groups or organizations).

    -   Elements 2, 3, 4 and 5 implies that there exist **interdependence** between the parties involved[^2]:

        -   Parties depend on each other to achieve their own preferred outcome [@lewicki_negociacion_2024, Chapter 1, p. 10]

    -   Element 6 considers that there are tangible and intangible aspects:

        -   **Tangibles**: aspects of which it is sought to reach an agreement within the negotiation (prices, terms of a contract, product specifications).

        -   **Intangibles**: psychological motivations implicit in a negotiation (personal values and emotions of the parties involved).

[^2]: If there is no interdependence, use a decision-making process other than negotiation

##

- Interdependence

    - By **interdependence** we refer to a situation in which *the parties need each other in order to accomplish their objectives and hence have the potential to influence each other* [@lewicki_negociacion_2024, p. 10]
    
        - Interdependence does not mean that everyone wants or needs the same thing. [@lewicki_negociacion_2024, p. 11]
        
        - Interdependence implies that you need to work together to accomplish something. [@lewicki_negociacion_2024, p. 11]

## {.smaller}

-   The type of interdependence that occurs between the parties affects the dynamics and the results of a negotiation:

    -   **Zero-sum or distributive situation**: for one party to obtain a profit,it is necessary for another party to obtain a loss.

    -   **Non-zero-sum or integrative situation**: there exists the ***possibility*** that all parties involved can obtain a profit without a party obtaining a loss.

-   The alternatives shape the interdependence where one way to analyze the alternatives is through the concept of:

    -   **Best Alternative to a Negotiated Agreement (BATAN)**[^3]

[^3]: For more details check out [@program_of_negotiation_batna_2012]

## {.smaller}

- Mutual adjustment

    -   Because the parties are interdependent during a negotiation there is a process called **mutual adjustment** where the parties try to influence one another to reach an acceptable agreement.
    
        -   At the beginning each party proposes an initial level on the tangible aspects of negotiation.
    
        -   Then there are replicas and concessions where the **bargaining range** tends to be reduced.
    
        -   If the **bargaining range** is reduced enough to become a point, an agreement is reached.
        
##

- Dilemmas in mutual adjustment

    - **Dilemma of honesty**: How much truth to tell the other parties? [@lewicki_negociacion_2024, p. 15]
    
        - You need to gain the trust of the other parties by telling the truth but not all the truth because you can not disclose all the information

    - **Dilemma of trust**: How much you should trust the other parties? [@lewicki_negociacion_2024, p. 15]
    
        - It depends on the reputation and past interactions with the other parties.

## 

- Approaches and tactics

    -   **Distributive Negotiation Approach**: it is conceived that there is only one winner in a particular situation and a course of action is sought to be the winner.
    
        -   Tactic: **claim value**[^4] by doing whatever is necessary to claim the reward or gain the largest piece possible [@lewicki_negociacion_2024, p. 16]
        
[^4]: For more details check out [@spangler_creating_2003]

##

- Approaches and tactics

    -   **Integrative Negotiation Approach**: attempts are made to find solutions so that both parties feel satisfied and reach their goals.
    
        -   Tactic: **create value**[^5] by finding the way for the parties involved to achieve their objectives by obtaining more resources

[^5]: For more details check out [@spangler_creating_2003]

# Conflict

## {.smaller}

-   For definitions of **conflict** please check out [@lewicki_negociacion_2024, p. 20]

    -   3 different definitions that are related are mentioned

-   For the different levels of **conflict** please check out [@lewicki_negociacion_2024, pp. 20-21]

    -   4 levels of conflict are mentioned

    -   We will not cover the first one in the course:

        -   *Intrapersonal or intrapsychic conflict*

-   For dysfunctions and functions of **conflict** please check out [@lewicki_negociacion_2024, pp. 21-22]

    -   Dysfunctions: negative aspects of conflict
    -   Functions: productive aspects of conflict

##

```{r}
#| label: fig-conflict-diagnostic-model
#| fig-cap: Conflict diagnostic model [@lewicki_negociacion_2024, p. 23] and [@greenhalgh_smr_1986, p. 47]

conflic_diag <- tibble(Dimension   = rep(
  c("Simmetry of conflict",
    "Involment of third parties",
    "Structure of the parties",
    "Interaction",
    "Interdependence",
    "Size of stakes",
    "Issue in question"), 
  times = 2),
  Level       = c(rep("Easy to resolve", times = 7),
                  rep("Difficult to resolve", times = 7)),
  label_text  = c("Equal damage",
                  "Trusted third parties",
                  "Strong Leadership",
                  "Long Term Relationship",
                  "Integrative situation",
                  "Small consequences",
                  "Divisible Issue",
                  "Asimmetry in damage suffered",
                  "No third parties",
                  "Disorganized",
                  "Single interaction",
                  "Distributive situation",
                  "Big consequences",
                  "Fundamental principles")) |> 
  mutate(Level = factor(Level, 
                        levels = c("Easy to resolve", 
                                   "Difficult to resolve")))

# Auxiliar data
conflic_diag_level <- conflic_diag |> 
  slice(1:7) |>
  select(-label_text) |> 
  set_names(nm = c("Dimension", "Level_easy")) |> 
  left_join(y = conflic_diag |> 
              slice(8:14) |>
              select(-label_text) |> 
              set_names(nm = c("Dimension", "Level_difficult")),
            by = join_by(Dimension))

conflic_diag |> 
  ggplot() +
  geom_segment(data = conflic_diag_level,
               aes(x = Level_easy, xend = Level_difficult,
                   y = Dimension,  yend = Dimension)) +
  geom_label(aes(x = Level, y = Dimension, 
                 label = label_text, fill = Dimension), 
             show.legend = FALSE,
             fontface    = "bold",
             size        = 3, 
             nudge_y     = 0.4,
             alpha       = 0.4) +
  scale_fill_manual(values = umng_palette[3:9]) +
  ggnewscale::new_scale(new_aes = "fill") +
  geom_point(aes(x = Level, y = Dimension, fill = Level),
             size  = 5,
             shape = 21, 
             color = "black") + 
  scale_fill_manual(values = umng_palette[1:2]) +
  labs(x = NULL) +
  theme(panel.border      = element_rect(fill = NA, 
                                         color = "black"),
        plot.background   = element_rect(fill = "white"),
        panel.background  = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white"), 
        axis.text.x       = element_blank(), 
        axis.ticks.x      = element_blank(),
        legend.position   = "bottom")
```

##

```{r}
#| label: fig-dual-concerns-model
#| fig-cap: Dual concerns model [@lewicki_negociacion_2024, p. 24] and [@pruitt_social_2004, p. 41][^6]

dual_concern_model <- tibble(x = c(1, 1, 3, 5, 5),
                             y = c(1, 5, 3, 1, 5),
                             labels = c("Inaction (No actuar)",
                                        "Yielding (Ceder)",
                                        "Compromising (Conciliar)",
                                        "Contending (Disputar)",
                                        "Problem solving (Resolver el problema)"))

dual_concern_model |> 
  ggplot() + 
  geom_label(aes(x = x, y = y, label = labels),
             size = 5,
             fill = "#fdc600") +
  geom_segment(data = tibble(x = 0,
                             y = 0,
                             yend = 0, xend = 6),
               aes(x = x, y = y,
                   xend = xend, yend = yend),
               color = "#043074") +
  geom_segment(data = tibble(x = 0,
                             y = 0,
                             yend = 6, xend = 0),
               aes(x = 0, y = 0,
                   yend = 6, xend = 0),
               color = "#043074") +
  scale_x_continuous(limits = c(0, 6)) +
  labs(x = "Concern about own outcomes",
       y = "Concern about other’s outcomes") +
  theme(panel.border      = element_rect(fill = NA, 
                                         color = "black"),
        plot.background   = element_rect(fill = "white"),
        panel.background  = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white"),
        legend.position   = "bottom", 
        axis.text         = element_blank(),
        axis.ticks        = element_blank())
```

[^6]: *Compromising* is not considered as a viable strategy by [@pruitt_social_2004] but [@lewicki_negociacion_2024] consider it

# Course and book structure

##

-  Cover in the course

    -   **Negotiation Fundamentals** 
    
        -  [@lewicki_negociacion_2024, chapters 1-5]
        
    -   **Negotiation Subprocesses**
    
        -  [@lewicki_negociacion_2024, chapters 6-8]
        
    -   **Concluding Comments**
    
        -  [@lewicki_negociacion_2024, chapter 20]

## {.smaller}

-  Not cover in the course
        
    -   **Negotiation Subprocesses**
    
        -  [@lewicki_negociacion_2024, chapter 9]
        
    -   **Negotiation Contexts**
    
        -  [@lewicki_negociacion_2024, chapters 10-13]
        
    -   **Individual Differences**
    
        -  [@lewicki_negociacion_2024, chapters 14-15]
        
    -   **Negotiation across Cultures**
    
        -  [@lewicki_negociacion_2024, chapter 16]

    -   **Resolving Differences**
    
        -  [@lewicki_negociacion_2024, chapter 17-19]

# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html) and [**$\pi$-thon**](https://www.python.org/about/)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://posit.co/download/rstudio-desktop/), [**Positron**](https://positron.posit.co/), [**Quarto**](https://quarto.org/) and the authors and maintainers of the packages [**tidyverse**](https://CRAN.R-project.org/package=tidyverse), [**ggnewscale**](https://CRAN.R-project.org/package=ggnewscale) and [**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

# References {.allowframebreaks}