---
title: Strategy and Tactics of Distributive Bargaining
author: 
  - Luis Francisco Gómez López
institute: 
  - FAEDIS
date: last-modified
format: 
  beamer:
    colortheme: dolphin
    fonttheme: structurebold
    theme: AnnArbor
    link-citations: true
    colorlinks: true
    toc: true
    slide-level: 2
    section-titles: false
    keep-tex: true
    include-in-header:
      file: 000_tex_files/preamble.tex
fig-cap-location: bottom
tbl-cap-location: top
knitr: 
  opts_chunk: 
    echo: false
    warning: false
    fig-align: center
    out-width: 85%
lang: en
bibliography: conflict_neg.bib
---

```{r}
#| label: libraries

library(tidyverse)
library(tidyquant)
library(ggrepel)
library(latex2exp)
```

# Please Read Me

##

-   Check the message **Welcome greeting** published in the News Bulletin Board.

-   Dear student please edit your profile uploading a photo where your face is clearly visible.

-   The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

-   This presentation is based on [@lewicki_negociacion_2024, Chapter 2]

# Purpose

##

Understand the basic elements of a distributive negotiation situation, as well as the strategies and tactics that can be applied in this context.

# Elements of a distributive negotiation situation

##

```{=tex}
\definecolor{red}{RGB}{227, 26, 28}
\definecolor{blue}{RGB}{44, 62, 80}
```

-   Buyer: Jackson ($\color{blue}{J}$) and Seller: Sofia ($\color{red}{S}$)[^1]

    -   Resistance point $\color{blue}{J}$, $\color{blue}{Rp_J}$: 150K USD
    -   Resistance point $\color{red}{S}$, $\color{red}{Rp_S}$: 130K USD

[^1]: Both $\color{blue}{J}$ and $\color{red}{S}$ have a common interest in cooperating but conflicting interests on how exactly to cooperate. For that reason a negotiation situation arises [@muthoo_bargaining_1999]

```{r}
#| label: fig-resistance-point
#| fig-cap: Resistance points
#| fig-height: 1.7
#| fig-width: 4.5

# Data
dist_neg1 <- tibble(x            = c(130, 150),
                    y            = rep(x = 0, times = 2),
                    negotiator   = c("Sofia", "Jackson"),
                    label_text   = c(TeX(r"($Rp_S$)"), TeX(r"($Rp_J$)")))

# Plot
plot_dist_neg1 <- dist_neg1 |> 
    ggplot() + 
    geom_segment(aes(x = 130, xend = 150, y = 0, yend = 0)) +
    geom_vline(aes(xintercept = x), 
               linetype = "dashed") +
    geom_point(aes(x = x, y = y, fill = negotiator),
               shape       = 21,
               size        = 5,
               color       = "black", 
               show.legend = FALSE) +
    geom_label_repel(aes(x = x, y = y, 
                         label = label_text, 
                         fill = negotiator),
                     color       = "white",
                     alpha       = 0.9, 
                     parse       = TRUE,
                     show.legend = FALSE) +
    scale_x_continuous(breaks = dist_neg1$x, 
                       labels = scales::label_number(prefix = "$", 
                                                     suffix = "K",
                                                     accuracy = 1)) +
    scale_fill_tq() +
    theme(panel.border      = element_rect(fill = NA, color = "black"),
          plot.background   = element_rect(fill = "#f3fcfc"),
          panel.background  = element_rect(fill = "#f3f7fc"), 
          panel.grid.major  = element_blank(),
          panel.grid.minor  = element_blank(),
          axis.title        = element_blank(),
          axis.text.x       = element_text(face = "bold", angle = 90, vjust = 0.5),
          axis.text.y       = element_blank(), 
          axis.ticks.y      = element_blank())

plot_dist_neg1 
```

##

```{=tex}
\definecolor{red}{RGB}{227, 26, 28}
\definecolor{blue}{RGB}{44, 62, 80}
```

-   Buyer: Jackson ($\color{blue}{J}$) and Seller: Sofia ($\color{red}{S}$)

    -   Initial offer $\color{blue}{J}$, $\color{blue}{Io_J}$: 133K USD
    -   Initial offer $\color{red}{S}$, $\color{red}{Io_S}$: 145K USD

```{r}
#| label: fig-initial-offer
#| fig-cap: Initial offers
#| fig-height: 1.7
#| fig-width: 4.5

# Data
dist_neg2 <- tibble(x            = c(130, 133, 145, 150),
                    y            = rep(x = 0, times = 4),
                    negotiator   = c("Sofia", "Jackson", "Sofia", "Jackson"),
                    label_text   = c(TeX(r"($Rp_S$)"),TeX(r"($Io_J$)"), 
                                     TeX(r"($Io_S$)"), TeX(r"($Rp_J$)")))

# Plot
plot_dist_neg2 <- dist_neg2 |> 
    ggplot() + 
    geom_segment(aes(x = 130, xend = 150, y = 0, yend = 0)) +
    geom_vline(aes(xintercept = x), 
               linetype = "dashed") +
    geom_point(aes(x = x, y = y, fill = negotiator),
               shape       = 21,
               size        = 5,
               color       = "black", 
               show.legend = FALSE) +
    geom_label_repel(aes(x = x, y = y, 
                         label = label_text, 
                         fill = negotiator),
                     color       = "white",
                     alpha       = 0.9,
                     parse       = TRUE,
                     show.legend = FALSE) +
    scale_x_continuous(breaks = dist_neg2$x, 
                       labels = scales::label_number(prefix = "$", 
                                                     suffix = "K", 
                                                     accuracy = 1)) +
    scale_fill_tq() +
    theme(panel.border      = element_rect(fill = NA, color = "black"),
          plot.background   = element_rect(fill = "#f3fcfc"),
          panel.background  = element_rect(fill = "#f3f7fc"), 
          panel.grid.major  = element_blank(),
          panel.grid.minor  = element_blank(),
          axis.title        = element_blank(),
          axis.text.x       = element_text(face = "bold", angle = 90, vjust = 0.5),
          axis.text.y       = element_blank(), 
          axis.ticks.y      = element_blank())

plot_dist_neg2 
```

##

```{=tex}
\definecolor{red}{RGB}{227, 26, 28}
\definecolor{blue}{RGB}{44, 62, 80}
```

-   Buyer: Jackson ($\color{blue}{J}$) and Seller: Sofia ($\color{red}{S}$)

    -   Target point $\color{blue}{J}$, $\color{blue}{Tp_J}$: 135K USD
    -   Target point $\color{red}{S}$, $\color{red}{Tp_S}$: 140K USD

```{r}
#| label: fig-target-point
#| fig-cap: Target points
#| fig-height: 1.7
#| fig-width: 4.5

# Data
dist_neg3 <- tibble(x            = c(130, 133, 135, 140, 145, 150),
                    y            = rep(x = 0, times = 6),
                    negotiator   = c("Sofia", "Jackson", "Jackson", "Sofia", "Sofia", "Jackson"),
                    label_text   = c(TeX(r"($Rp_S$)"),TeX(r"($Io_J$)"),
                                     TeX(r"($Tp_J$)"), 
                                     TeX(r"($Tp_S$)"),TeX(r"($Io_S$)"), 
                                     TeX(r"($Rp_J$)")))

# Plot
plot_dist_neg3 <- dist_neg3 |> 
    ggplot() + 
    geom_segment(aes(x = 130, xend = 150, y = 0, yend = 0)) +
    geom_vline(aes(xintercept = x), 
               linetype = "dashed") +
    geom_point(aes(x = x, y = y, fill = negotiator),
               shape       = 21,
               size        = 5,
               color       = "black", 
               show.legend = FALSE) +
    geom_label_repel(aes(x = x, y = y, 
                         label = label_text, 
                         fill = negotiator),
                     color       = "white",
                     alpha       = 0.9, 
                     parse       = TRUE, 
                     show.legend = FALSE) +
    scale_x_continuous(breaks = dist_neg3$x, 
                       labels = scales::label_number(prefix = "$", 
                                                     suffix = "K", 
                                                     accuracy = 1)) +
    scale_fill_tq() +
    theme(panel.border      = element_rect(fill = NA, color = "black"),
          plot.background   = element_rect(fill = "#f3fcfc"),
          panel.background  = element_rect(fill = "#f3f7fc"), 
          panel.grid.major  = element_blank(),
          panel.grid.minor  = element_blank(),
          axis.title        = element_blank(),
          axis.text.x       = element_text(face = "bold", angle = 90, vjust = 0.5),
          axis.text.y       = element_blank(), 
          axis.ticks.y      = element_blank())

plot_dist_neg3
```

##

```{=tex}
\definecolor{red}{RGB}{227, 26, 28}
\definecolor{blue}{RGB}{44, 62, 80}
```

-   Buyer: Jackson ($\color{blue}{J}$) and Seller: Sofia ($\color{red}{S}$)

    -   BATNA $\color{blue}{J}$, $\color{blue}{BATNA_J}$: 142K USD
    -   BATNA $\color{red}{S}$, $\color{red}{BATNA_S}$: 134K USD

```{r}
#| label: fig-batna
#| fig-cap: Best alternatives to a negotiated agreement 
#| fig-height: 1.7
#| fig-width: 4.5
# Data
dist_neg4 <- tibble(x            = c(130, 133, 134, 135, 140, 142, 145, 150),
                    y            = rep(x = 0, times = 8),
                    negotiator   = c("Sofia", "Jackson", "Sofia", "Jackson", "Sofia", "Jackson", "Sofia", "Jackson"),
                    label_text   = c(TeX(r"($Rp_S$)"),TeX(r"($Io_J$)"),
                                     TeX(r"($BATNA_S$)"), TeX(r"($Tp_J$)"), 
                                     TeX(r"($Tp_S$)"), TeX(r"($BATNA_J$)"),
                                     TeX(r"($Io_S$)"), 
                                     TeX(r"($Rp_J$)")))

# Plot
plot_dist_neg4 <- dist_neg4 |> 
    ggplot() + 
    geom_segment(aes(x = 130, xend = 150, y = 0, yend = 0)) +
    geom_vline(aes(xintercept = x), 
               linetype = "dashed") +
    geom_point(aes(x = x, y = y, fill = negotiator),
               shape       = 21,
               size        = 5,
               color       = "black", 
               show.legend = FALSE) +
    geom_label_repel(aes(x = x, y = y, 
                         label = label_text, 
                         fill = negotiator),
                     color       = "white",
                     alpha       = 0.9,
                     parse       = TRUE,
                     show.legend = FALSE) +
    scale_x_continuous(breaks = dist_neg4$x, 
                       labels = scales::label_number(prefix = "$", 
                                                     suffix = "K", 
                                                     accuracy = 1)) +
    scale_fill_tq() +
    theme(panel.border      = element_rect(fill = NA, color = "black"),
          plot.background   = element_rect(fill = "#f3fcfc"),
          panel.background  = element_rect(fill = "#f3f7fc"), 
          panel.grid.major  = element_blank(),
          panel.grid.minor  = element_blank(),
          axis.title        = element_blank(),
          axis.text.x       = element_text(face = "bold", angle = 90, vjust = 0.5),
          axis.text.y       = element_blank(), 
          axis.ticks.y      = element_blank())

plot_dist_neg4
```

##

```{=tex}
\definecolor{red}{RGB}{227, 26, 28}
\definecolor{blue}{RGB}{44, 62, 80}
```

-   Buyer: Jackson ($\color{blue}{J}$) and Seller: Sofia ($\color{red}{S}$)

    -   Bargaining range[^2]: $[\text{\$130K},\text{\$150K}]$

        -   Is the result of $[\text{\$130K}, \infty) \cap [0, \text{\$150K}]$
        -   If the bargaining range is $\neq \emptyset$, $Rp_S \leq BATNA_J$ and $BATNA_S \leq Rp_J$ then an aggrement between $\color{blue}{J}$ and $\color{red}{S}$ is possible

[^2]: Also known as settlement range or zone of potential agreement

```{r}
#| label: fig-bargaining-range
#| fig-cap: Bargaining range 
#| fig-height: 1.7
#| fig-width: 4.5

plot_dist_neg4
```

##

```{=tex}
\definecolor{red}{RGB}{227, 26, 28}
\definecolor{blue}{RGB}{44, 62, 80}
```

-   Buyer: Jackson ($\color{blue}{J}$) and Seller: Sofia ($\color{red}{S}$)

    -   In this example the settlement point $\in [Tp_J, Tp_S] = [\text{\$135K}, \text{\$140K}]$

```{r}
#| label: fig-settlement-point
#| fig-cap: Possible values for a settlement point 
#| fig-height: 1.7
#| fig-width: 4.5

plot_dist_neg4
```

##

```{=tex}
\definecolor{red}{RGB}{227, 26, 28}
\definecolor{blue}{RGB}{44, 62, 80}
```

-   In the case of a distributive negotiation, it is most likely that the settlement point does not correspond to the target points of the participants in the negotiation.

-   In the example of $\color{blue}{J}$ and $\color{red}{S}$ it was implicitly assumed that the negotiation element was the price. This element is known as a **bargaining mix**[^3] where it does not necessarily have to be a single element.

    -   For example $\color{blue}{J}$ and $\color{red}{S}$ also can include in the bargaining mix the final date of the sale, whether or not renovations are included or if other items such as furniture and household appliances are added.

[^3]: It is define as the set of issues that will be negotiated

# Strategies

##

```{dot}
//| label: fig-strategies-distributive-bargaining
//| fig-cap: Strategies distributive bargaining [@lewicki_negociacion_2024, pp. 39-41]
//| fig-width: 5.5
//| fig-height: 2.5

digraph boxes_and_circles {

  graph [rankdir = LR]

  node [shape = circle, color = "#2C3E50"]
 
  "Strategies";

  node [shape = oval, color = "#E31A1C"]

  "Discover\nor\nestimate\nthe\nresistance point";
  "Influence\nthe\nresistance point\nof the\ncounterpart";
  
  node [shape = oval, color = "#18BC9C"]
  
  "4 major\npropositions";
  
  "Strategies" -> "Discover\nor\nestimate\nthe\nresistance point";
  "Strategies" -> "Influence\nthe\nresistance point\nof the\ncounterpart";
  
  "Influence\nthe\nresistance point\nof the\ncounterpart" -> "4 major\npropositions";
  
}
```

# Tactics

##

```{dot}
//| label: fig-tactics-distributive-bargaining-1
//| fig-cap: Tactics distributive bargaining [@lewicki_negociacion_2024, pp. 42-48]
//| fig-width: 4.5
//| fig-height: 2.5

digraph boxes_and_circles {

  graph [rankdir = LR]

  node [shape = circle, color = "#2C3E50"]
 
  "Tactics";

  node [shape = circle, color = "#E31A1C"]

  "Assess the other\nparty's";
  "Manage the other\nparty's impressions";
  
  node [shape = oval, color = "#18BC9C"]

  "Indirect\nassessment";
  "Direct\nassessment";
  
  "Screening\nactivities";
  "Direct action to\nalter impressions";
  
  node [shape = oval, color = "#CCBE93"]
  
  "Target\npoints";
  "Resistance\npoints";
  "Costs of\nending\nthe negotiation";
  
  "Tactics" -> "Assess the other\nparty's";
  "Tactics" -> "Manage the other\nparty's impressions";
  
  "Assess the other\nparty's" -> "Indirect\nassessment";
  "Assess the other\nparty's" -> "Direct\nassessment";
  "Manage the other\nparty's impressions" -> "Screening\nactivities";
  "Manage the other\nparty's impressions" -> "Direct action to\nalter impressions";
  
  "Indirect\nassessment" -> "Target\npoints";
  "Indirect\nassessment" -> "Resistance\npoints";
  "Indirect\nassessment" -> "Costs of\nending\nthe negotiation";
  "Direct\nassessment" -> "Target\npoints";
  "Direct\nassessment" -> "Resistance\npoints";
  "Direct\nassessment" -> "Costs of\nending\nthe negotiation";
  
  "Screening\nactivities" -> "Target\npoints";
  "Screening\nactivities" -> "Resistance\npoints";
  "Screening\nactivities" -> "Costs of\nending\nthe negotiation";
  "Direct action to\nalter impressions" -> "Target\npoints";
  "Direct action to\nalter impressions" -> "Resistance\npoints";
  "Direct action to\nalter impressions" -> "Costs of\nending\nthe negotiation";
  
}
```

##

```{dot}
//| label: fig-tactics-distributive-bargaining-2
//| fig-cap: Tactics distributive bargaining [@lewicki_negociacion_2024, pp. 42-48]
//| fig-width: 4.5
//| fig-height: 2.5

digraph boxes_and_circles {

  graph [rankdir = LR]

  node [shape = circle, color = "#2C3E50"]
 
  "Tactics";

  node [shape = circle, color = "#E31A1C"]

  "Modify the\nperception";
  "Increase the costs\nof delaying negotiations";
  
  node [shape = oval, color = "#18BC9C"]
  
  "Alter the impressions\nof the counterpart";
  "Make demands and\npositions appear\nmore or less unattractive";
  "Disruptive action";
  "Alliance with outsiders";
  "Schedule manipulation";
  
  node [shape = oval, color = "#CCBE93"]
  
  "Target\npoints";
  "Resistance\npoints";
  "Costs of\nending\nthe negotiation";
  
  "Tactics" -> "Modify the\nperception";
  "Tactics" -> "Increase the costs\nof delaying negotiations";
  
  "Modify the\nperception" -> "Alter the impressions\nof the counterpart";
  "Modify the\nperception" -> "Make demands and\npositions appear\nmore or less unattractive";
  "Increase the costs\nof delaying negotiations" -> "Disruptive action";
  "Increase the costs\nof delaying negotiations" -> "Alliance with outsiders";
  "Increase the costs\nof delaying negotiations" -> "Schedule manipulation";

  "Alter the impressions\nof the counterpart" -> "Target\npoints";
  "Alter the impressions\nof the counterpart" -> "Resistance\npoints";
  "Alter the impressions\nof the counterpart" -> "Costs of\nending\nthe negotiation";
  "Make demands and\npositions appear\nmore or less unattractive" -> "Target\npoints";
  "Make demands and\npositions appear\nmore or less unattractive" -> "Resistance\npoints";
  "Make demands and\npositions appear\nmore or less unattractive" -> "Costs of\nending\nthe negotiation";
  
}
```

# Aspects about taking positions during a negotiation

##

-   In [@lewicki_negociacion_2024, Chapter 2, p 48-55] the recommended positions to be taken with regard to the following elements are pointed out:

    -   **Opening offers**

        -   Who should make the first offer?
        -   What should be the initial offer?
        -   Should the offer be perceived as low, moderate or high by the other party?
        -   Should the initial offer be near or far from our own resistance point?

    -   **Opening stance**

        -   Should a moderate or aggressive stance be taken?

    -   **Initial concessions**

        -   How wide should the initial concession be?

    -   **Final offer**

        -   How to communicate that a particular offer is a final offer?

# Commitments

##

-  **Commitment** means taking a stance in a negotiation and making a explicit or implicit promise about future actions.

    -  A commitment is a way to create a bargaining position by specifying a future action if a position is not reached.
    
    -  A commitment aims to clarify the negotiator's planned actions and eliminate any uncertainty about their intentions.
    
        -  However, they may also fix a negotiator to a particular position [@lewicki_negociacion_2024, Chapter 2, p 56]
        -  That is why when making commitments, one should also make contingency plans for a graceful exit if needed
        
        -  Also it is important to point out that good, sound and deliberate commitments take time to establish so it is important to don't commit prematurely [@lewicki_negociacion_2024, Chapter 2, p 56]
        
##        

```{dot}
//| label: fig-ways-to-abandon-committed-position
//| fig-cap: Ways to abandon a committed position [@lewicki_negociacion_2024, pp. 58-59]
//| fig-width: 4.5
//| fig-height: 2.5

digraph boxes_and_circles {

  graph [rankdir = TB]

  node [shape = circle, color = "#2C3E50"]
 
  "Ways to Abandon a\nCommitted Position";

  node [shape = circle, color = "#E31A1C"]

  "Plan a way out";
  "Let it die silently";
  "Restate the Commitment";
  "Minimize the Damage";
  
  "Ways to Abandon a\nCommitted Position" -> "Plan a way out";
  "Ways to Abandon a\nCommitted Position" -> "Let it die silently";
  "Ways to Abandon a\nCommitted Position" -> "Restate the Commitment";
  "Ways to Abandon a\nCommitted Position" -> "Minimize the Damage";
  
}
```

# Closing a deal

##

-   The negotiations seek to reach an agreement if possible[^4]. In order to achieve that goal with a greater probability, the following practices are recommended based on what is pointed out in [@lewicki_negociacion_2024, Chapter 2, p 59-60]:

    -   Offer similar options to the other parties to make the negotiation more flexible.

    -   Assume a closing technique or stance.

    -   Split the difference when a **mutual adjustment** process has been carried out.

    -   Use **exploding offers** that refers to setting a deadline where a specific proposal is in force for a limited time.

    -   Use **sweeteners** that refers to granting special concessions before closing.

[^4]: Remember that negotiation as a form of decision making is not the only method that exists!

# How to face hardball tactics?

##

-   Sometimes in negotiations people use **hardball tactics**. In general in [@lewicki_negociacion_2024, Chapter 2, p 61] the authors recommend not to use them as they cause damage in the negotiation process. However, it is important to know them and know how to deal with them.

-   In [@lewicki_negociacion_2024, Chapter 2, pp 61-62] 4 possible strategies to face these tactics are pointed out:

    -  Discuss them
    -  Ignore them
    -  Respond in kind
    -  Co-opt the other party
    
## 

```{dot}
//| label: fig-hard-ball-tactics
//| fig-cap: Typical hardball tactics [@lewicki_negociacion_2024, p 63]
//| fig-width: 4.5
//| fig-height: 2.5

digraph boxes_and_circles {

  graph [rankdir = TB]

  node [shape = circle, color = "#2C3E50"]
 
  "Hardball tactics";

  node [shape = circle, color = "#E31A1C"]

  "Good cop/bad cop";
  "Lowball/highball";
  "Bogey";
  "Chicken";
  "Intimidation";
  "Aggressive\nbehavior";
  "Snow job";
  
  "Hardball tactics" -> "Good cop/bad cop";
  "Hardball tactics" -> "Lowball/highball";
  "Hardball tactics" -> "Bogey";
  "Hardball tactics" -> "Chicken";
  "Hardball tactics" -> "Intimidation";
  "Hardball tactics" -> "Aggressive\nbehavior";
  "Hardball tactics" -> "Snow job";

}
```

# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html) and [**$\pi$-thon**](https://www.python.org/about/)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://rstudio.com/products/rstudio/), [**Quarto**](https://quarto.org/) and the authors and maintainers of the packages [**tidyverse**](https://CRAN.R-project.org/package=tidyverse), [**tidyquant**](https://CRAN.R-project.org/package=tidyquant), [**ggrepel**](https://CRAN.R-project.org/package=ggrepel), [**DiagrammeR**](https://CRAN.R-project.org/package=DiagrammeR) and [**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

-   To the [**Linux kernel community**](https://www.kernel.org/category/about.html) for allowing me the possibility to use some [**Linux distributions**](https://static.lwn.net/Distributions/) as my main [**OS**](https://en.wikipedia.org/wiki/Operating_system) without paying for a license

# References {.allowframebreaks}
