---
title: "Negotiation: Strategy and Planning"
author: "Luis Francisco Gomez Lopez"
date: 2023-07-14
format:
  beamer:
    colortheme: dolphin
    fonttheme: structurebold
    theme: AnnArbor
    link-citations: true
    linkcolor: blue
    include-in-header:
      - text: |
          \usepackage{booktabs}
          \usepackage{longtable}
          \usepackage{array}
          \usepackage{multirow}
          \usepackage{wrapfig}
          \usepackage{float}
          \usepackage{colortbl}
          \usepackage{pdflscape}
          \usepackage{tabu}
          \usepackage{threeparttable}
          \usepackage{threeparttablex}
          \usepackage[normalem]{ulem}
          \usepackage{makecell}
          \usepackage{xcolor}
          \usepackage{fontawesome5}
bibliography: conflict_neg.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      message   = FALSE,
                      fig.align = "center")
```

```{r libraries}

library(DiagrammeR)
library(tufte)
```

# Contents

- Please Read Me
- Purpose
- The importance of planning in negotiations
- Key steps in the planning process
- Goals
- Strategies
- Planning process
- Acknowledgments
- References

# Please Read Me

- Check the message __Welcome greeting__ published in the News Bulletin Board.

- Dear student please edit your profile uploading a photo where your face is clearly visible.

- The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

- This presentation is based on [@lewicki_essentials_2016, Chapter 4]

# Purpose

Understand and explore the main elements of the negotiation strategy, the process to select a strategy as well as the way in which most negotiations evolve to effectively plan a negotiation.

# The importance of planning in negotiations

- Some quotes about planning:

> "Everyone has a plan until they get punched in the mouth"
>
> `r tufte::quote_footer('--- Mike Tyson')`

> "No battle plan ever survives contact with the enemy"
>
> `r tufte::quote_footer('--- Helmuth von Moltke the Elder')`

- Therefore, there is no need to plan? 

    + Absolutely not, but if your plan is not flexible you will get punched in the mouth and no part of your plan will survive when you negotiate with the counterpart.
    
    + Also don't expect everything will work millimetrically according to the plan because the counterpart might also have a plan and they will react strategically to your plan.
    
# The importance of planning in negotiations

- Without planning results occur more by chance than by negotiator effort

    + Behaviour of Successful Negotiators: Find some successful  negotiators and watch them  during  actual  negotiations  to  find  how  they  do  it.
        
        + [@rackham_effective_1978]: planning is the foundation for any successful negotiation
        + [@rackham_effective_1978-1]: exploration of options, common ground, long-term  implications, setting limits[^1] and use an issue  planning method over sequence planning[^2]

- Also check out in the __Links of interest__ the video: How to negotiate properly? [@magic_markers_como_2018]

[^1]: Plan in terms of a range and not around a fixed point
[^2]: Issues are independent and not linked by a sequence

# Key steps in the planning process

- Relationship between key steps in the planning Process [@lewicki_essentials_2016, Chapter 4, p 90]

```{r, out.width="90%", dev='png'}

grViz("
digraph boxes_and_circles {

  graph [rankdir = LR]

  node [shape = circle, color = '#2C3E50']
 
  'Goals';

  node [shape = circle, color = '#E31A1C']

  'Strategy';
  
  node [shape = circle, color = '#18BC9C']
  
  'Planning';
  
  'Goals' -> 'Strategy';
  
  'Strategy' -> 'Planning'
}
")
```

# Goals 

- The first step to execute a negotiation strategy is to determine one’s goals.

    + __Substantive__: goals about tangible factors (price, terms of a contract, product specifications)
    + __Intangible__: goals about intangible factors (personal values and emotions of the parties such as defeating the other party or reaching a conciliation at any cost)
    + __Procedural__: goals about how the negotiation process occurs (define plans or just have a voice during the negotiation) 
    
- The negotiator must identify what kind of goals to pursue[^3]. What definitely cannot be ignored is the substantive goals given that it refers to the tangible aspects.

[^3]: Identifying goals in practical terms simply means making a list to prioritize them

# Goals 

- When defining goals take into account these 4 aspects [@lewicki_essentials_2016, Chapter 4, p 90-91]:

    + Goals are specific objectives that are sought realistically.
    + Own goals can potentially be linked to the goals of the other parties.[^4]
    + If in identifying the goals these are not attainable then:
    
        + Modify them so that they are attainable
        + Discard the negotiation as an option to reach an agreement[^5]
        
    + The goals that are identified must be concrete, specific and measurable so that it is easier to:
    
        + Communicate to the other party what you want to achieve
        + Understand what the other parties want
        + Determine if the proposed proposals satisfy what you want to achieve
        
[^4]:It is easier to reach an agreement if you have common goals
[^5]: Remember that negotiation as a form of __decision making__ is not the only method that exists! 

# Strategies

- Different engagement strategies in a negotiation context[^6]

```{r, out.width="85%", dev='png'}

grViz("
digraph boxes_and_circles {

  graph [rankdir = TB]

  node [shape = circle, color = '#2C3E50']
 
  'Strategy';

  node [shape = circle, color = '#E31A1C']

  'Competitive\n(Distributive context)';
  'Collaborative\n(Integrative context)';
  'Accomodative';
  
  'Strategy' -> 'Competitive\n(Distributive context)';
  'Strategy' -> 'Collaborative\n(Integrative context)';
  'Strategy' -> 'Accomodative';
}
")
```

[^6]: Check out [@lewicki_essentials_2016, Chapter 4, p 94, Table 4.1] for more details

# Planning process

- Key steps in the planning process [@lewicki_essentials_2016, Chapter 4, p 96-112]

```{r, out.width="90%", dev='png'}

grViz("
digraph boxes_and_circles {

  graph [rankdir = LR]

  node [shape = circle, color = '#2C3E50']
 
  'Planning';

  node [shape = oval, color = '#E31A1C']

  'Defining the\nnegotiating goal';
  'List of major issues';
  'Organize Issues';
  'Defining the Interests';
  'Knowing Your BATNA'
  
  node [shape = oval, color = '#18BC9C']
  
  'Assembling';
  'Prioritization';
  'Substantive';
  'Process-based';
  'Relationship-based';
  
  node [shape = oval, color = '#CCBE93']
  
  'Ranking';
  'Defining the\nBargaining Mix';
  
  'Planning' -> 'Defining the\nnegotiating goal';
  'Planning' -> 'List of major issues';
  'Planning' -> 'Organize Issues';
  'Planning' -> 'Defining the Interests';
  'Planning' -> 'Knowing Your BATNA'
  
  'Organize Issues' -> 'Assembling';
  'Organize Issues' -> 'Prioritization';
  'Defining the Interests' -> 'Substantive';
  'Defining the Interests' -> 'Process-based';
  'Defining the Interests' -> 'Relationship-based';
  
  'Prioritization' -> 'Ranking';
  'Prioritization' -> 'Defining the\nBargaining Mix';
  
}
")
```

# Planning process

- Key steps in the planning process [@lewicki_essentials_2016, Chapter 4, p 96-112]

```{r, out.width="90%", dev='png'}

grViz("
digraph boxes_and_circles {

  graph [rankdir = LR]

  node [shape = circle, color = '#2C3E50']
 
  'Planning';

  node [shape = oval, color = '#E31A1C']

  'Knowing your\nresistance point';
  'Gathering information\nabout the other party';
  'Setting one’s own targets\nand opening bids';
  'Assessing the social\ncontext of negotiation';
  'Presenting the issues\nto the other Party'
  
  node [shape = oval, color = '#18BC9C']
  
  'Direct actors';
  'Opposition actors';
  'Indirect actors';
  'Interested observers';
  'Environmental factors';
  
  node [shape = oval, color = '#CCBE93']
  
  'Constituents';
  
  'Planning' -> 'Knowing your\nresistance point';
  'Planning' -> 'Gathering information\nabout the other party';
  'Planning' -> 'Setting one’s own targets\nand opening bids';
  'Planning' -> 'Assessing the social\ncontext of negotiation';
  'Planning' -> 'Presenting the issues\nto the other Party'
  
  'Assessing the social\ncontext of negotiation' -> 'Direct actors';
  'Assessing the social\ncontext of negotiation' -> 'Opposition actors';
  'Assessing the social\ncontext of negotiation' -> 'Indirect actors';
  'Assessing the social\ncontext of negotiation' -> 'Interested observers';
  'Assessing the social\ncontext of negotiation' -> 'Environmental factors';
  
  'Interested observers' -> 'Constituents';
  
}
")
```

# Acknowledgments

- To my family that supports me

- To the taxpayers of Colombia and the __[UMNG students](https://www.umng.edu.co/estudiante)__ who pay my salary

- To the __[Business Science](https://www.business-science.io/)__ and __[R4DS Online Learning](https://www.rfordatasci.com/)__ communities where I learn __[R](https://www.r-project.org/about.html)__ 

- To the __[R Core Team](https://www.r-project.org/contributors.html)__, the creators of __[RStudio IDE](https://rstudio.com/products/rstudio/)__ and the authors and maintainers of the packages __[DiagrammeR](https://CRAN.R-project.org/package=DiagrammeR)__, __[tufte](https://CRAN.R-project.org/package=tufte)__, and __[tinytex](https://CRAN.R-project.org/package=tinytex)__ for allowing me to access these tools without paying for a license

- To the __[Linux kernel community](https://www.kernel.org/category/about.html)__ for allowing me the possibility to use some __[Linux distributions](https://static.lwn.net/Distributions/)__ as my main __[OS](https://en.wikipedia.org/wiki/Operating_system)__ without paying for a license

# References