---
title: "Negotiation: Strategy and Planning"
subtitle: Session 1
author:
  - name: 
      given: Luis F.
      family: Gómez L.
    orcid: 0000-0002-2877-9015
    email: luis.gomezl@unimilitar.edu.co
    affiliations:
      - name: Distance Learning Faculty
toc: true
toc-depth: 1
slide-level: 2
format:
  presentation-umng-revealjs:
    logo-vertical-strip: 000_logos/logo-white-vertical.png 
    embed-resources: true
    mermaid-format: png
    reference-location: document
date: last-modified
date-format: "D MMMM, YYYY"
knitr: 
  opts_chunk: 
    echo: false
    warning: false
    fig-align: center
    out-width: 85%
lang: en
bibliography: conflict_neg.bib
csl: apa-7.csl
editor: source
---

```{r}
#| label: libraries

library(tufte)
```

```{r}
#| label: palette

umng_palette <- c(
  "#043074",
  "#fdc600",
  "#0d5e30",
  "#ee2a24",
  "#fc6700",
  "#00b3f0",
  "#6e3c1a",
  "#f8941c",
  "#8e44ad",
  "#2c3e50",
  "#16a085",
  "#c0392b",
  "#e91e63"
)
```

# Please Read Me

##

-   Check the message **Welcome greeting** published in the News Bulletin Board.

-   Dear student please edit your profile uploading a photo where your face is clearly visible.

-   The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

-   This presentation is based on [@lewicki_negociacion_2024, Chapter 4]

##

- **Purpose**

    - Understand and explore the main elements of the negotiation strategy, the process to select a strategy as well as the way in which most negotiations evolve to effectively plan a negotiation.

# The importance of planning in negotiations

## {.smaller}

-   Some quotes about planning:

> "Everyone has a plan until they get punched in the mouth"

> `r quote_footer("--- Mike Tyson")`

> "No battle plan ever survives contact with the enemy"

> `r quote_footer("--- Helmuth von Moltke the Elder")`

-   Therefore, there is no need to plan?

    -   Absolutely not, but if your plan is not flexible you will get punched in the mouth and no part of your plan will survive when you negotiate with the counterpart.

    -   Also don"t expect everything will work millimetrically according to the plan because the counterpart might also have a plan and they will react strategically to your plan.
    
## {.smaller}

-   Without planning results occur more by chance than by negotiator effort

    -   Behaviour of Successful Negotiators: Find some successful negotiators and watch them during actual negotiations to find how they do it.

        -   [@rackham_effective_1978]: planning is the foundation for any successful negotiation
        -   [@rackham_effective_1978-1]: exploration of options, common ground, long-term implications, setting limits[^1] and use an issue planning method over sequence planning[^2]

-   Also check out in the **Links of interest** the video: How to negotiate properly? [@magic_markers_como_2018]

[^1]: Plan in terms of a range and not around a fixed point

[^2]: Issues are independent and not linked by a sequence

# Key steps in the planning process

## {.center}

```{mermaid}
%%| label: fig-relation-ship-key-steps-planning-process
%%| fig-cap: Relationship between key steps in the planning process [@lewicki_negociacion_2024, Chapter 4, p. 105]
%%| fig-height: 2

flowchart LR
    A("Goals") --> A1("Strategy")
    A1 --> A11("Planning")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11 UmngNode;
```

# Goals

## {.smaller}

-   The first step to execute a negotiation strategy is to determine one’s goals.

    -   **Substantive**: goals about tangible factors (price, terms of a contract, product specifications)
    -   **Intangible**: goals about intangible factors (personal values and emotions of the parties such as defeating the other party or reaching a conciliation at any cost)
    -   **Procedural**: goals about how the negotiation process occurs (define plans or just have a voice during the negotiation)

-   The negotiator must identify what kind of goals to pursue[^3]. What definitely cannot be ignored is the substantive goals given that it refers to the tangible aspects.

[^3]: Identifying goals in practical terms simply means making a list to prioritize them

## {.smaller}

-   When defining goals take into account these 4 aspects [@lewicki_negociacion_2024, Chapter 4, p 105-106]:

    -   Goals are specific objectives that are sought realistically.

    -   Own goals can potentially be linked to the goals of the other parties.[^4]

    -   If in identifying the goals these are not attainable then:

        -   Modify them so that they are attainable
        -   Discard the negotiation as an option to reach an agreement[^5]

    -   The goals that are identified must be concrete, specific and measurable so that it is easier to:

        -   Communicate to the other party what you want to achieve
        -   Understand what the other parties want
        -   Determine if the proposed proposals satisfy what you want to achieve

[^4]: It is easier to reach an agreement if you have common goals

[^5]: Remember that negotiation as a form of **decision making** is not the only method that exists!

# Strategies

##

```{mermaid}
%%| label: fig-engagement-strategies
%%| fig-cap: Different engagement strategies in a negotiation context  [@lewicki_negociacion_2024, Chapter 4, p. 110]
%%| fig-height: 4

flowchart TB
    A("Strategy") --> A1("`Competitive
    **(Distributive context)**`")
    A --> A2("`Collaborative
    **(Integrative context)**`")
    A --> A3("Accomodative")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A2,A3 UmngNode;
```

# Phases of negotiation

##

-  Before exploring the specific planning processes for negotiation, it is important to understand the typical phases in a negotiation

    -  7 key steps to an ideal negotiation process [@greenhalgh_managing_2001, p. 164]:

        -  Preparation
        -  Relationship building
        -  Information gathering
        -  Information using
        -  Bidding
        -  Closing the deal
        -  Implementing the agreement

# Planning process

##

```{mermaid}
%%| label: fig-key-steps-planning-process-1
%%| fig-cap: Key steps in the planning process [@lewicki_negociacion_2024, Chapter 4, pp. 114-128]
%%| fig-height: 5

flowchart TB
    A("Planning") --> A1("Defining the negotiating goal")
    A --> A2("List the major issues related to achieving the goal")
    A --> A3("Organize Issues")
    A --> A4("Defining the Interests")
    A --> A5("Knowing Your BATNA")
    A3 --> A31("Assembling")
    A3 --> A32("Prioritization")
    A4 --> A41("Substantive")
    A4 --> A42("Process-based")
    A4 --> A43("Relationship-based")
    A32 --> A321("Ranking")
    A32 --> A322("Defining the Bargaining Mix")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A2,A3,A31,A32,A321,A322,A4,A41,A42,A43,A5 UmngNode;
```

##

```{mermaid}
%%| label: fig-key-steps-planning-process-2
%%| fig-cap: Key steps in the planning process [@lewicki_negociacion_2024, Chapter 4, pp. 114-128]
%%| fig-height: 5

flowchart TB
    A("Planning") --> A1("Knowing your resistance point")
    A --> A2("Gathering information about the other party")
    A --> A3("Setting one’s own targets and opening bids")
    A --> A4("Assessing the social context of negotiation")
    A --> A5("Presenting the issues to the other Party")
    A4 --> A41("Direct actors")
    A4 --> A42("Opposition actors")
    A4 --> A43("Indirect actors")
    A4 --> A44("Interested observers")
    A4 --> A45("Environmental factors")
    A44 --> A441("Constituents")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A2,A3,A4,A41,A42,A43,A44,A441,A45,A5 UmngNode;
```

# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html) and [**$\pi$-thon**](https://www.python.org/about/)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://rstudio.com/products/rstudio/), [**Quarto**](https://quarto.org/) and the authors and maintainers of the packages [**tufte**](https://CRAN.R-project.org/package=tufte) and [**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

# References {.allowframebreaks}