---
title: Perception, Cognition, and Emotion
subtitle: Session 1
author:
  - name: 
      given: Luis F.
      family: Gómez L.
    orcid: 0000-0002-2877-9015
    email: luis.gomezl@unimilitar.edu.co
    affiliations:
      - name: Distance Learning Faculty
toc: true
toc-depth: 1
slide-level: 2
format:
  presentation-umng-revealjs:
    logo-vertical-strip: 000_logos/logo-white-vertical.png 
    embed-resources: true
    mermaid-format: png
    reference-location: document
date: last-modified
date-format: "D MMMM, YYYY"
knitr: 
  opts_chunk: 
    echo: false
    warning: false
    fig-align: center
    out-width: 85%
lang: en
bibliography: conflict_neg.bib
csl: apa-7.csl
editor: source
---

```{r}
#| label: libraries


```

```{r}
#| label: palette

umng_palette <- c(
  "#043074",
  "#fdc600",
  "#0d5e30",
  "#ee2a24",
  "#fc6700",
  "#00b3f0",
  "#6e3c1a",
  "#f8941c",
  "#8e44ad",
  "#2c3e50",
  "#16a085",
  "#c0392b",
  "#e91e63"
)
```

# Please Read Me

##

-   Check the message **Welcome greeting** published in the News Bulletin Board.

-   Dear student please edit your profile uploading a photo where your face is clearly visible.

-   The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

-   This presentation is based on [@lewicki_negociacion_2024, Chapter 6]

##

- **Purpose**

    - Understand how perceptions, knowledge and emotions affect a negotiation process.

# Perception

##

-   **Perception** is defined as the process by which individuals acquire information from their environment through the senses.

    -   Individuals actively construct impressions based on this information, guiding their subsequent behavior

-   The distortions of the perception may occur when the impressions that are generated don't reflect the real aspects of a situation.

    -   This type of distortions are known as **cognitive biases**

##

-   Example of a distortion in perception

![Checker Shadow illusion [@adelson_checkershadow_1995]](000_images/005_checkershadow_double_full.jpg){.nostretch #fig-checker-shadow-illusion width=90% fig-align="center"}

##

-   Example of a distortion in perception

![Sander's parallelogram [@luckiesh_visual_2017]](000_images/005_sander_Illusion.png){.nostretch #fig-sanders-parallelogram width=90% fig-align="center"}

##

-   Example of a distortion in perception

![Müller-Lyer Illusion [@zeman_muller-lyer_2013]](000_images/005_muller_lLyer_circle.png){.nostretch #fig-muller-lyer-illusion width=50% fig-align="center"}

# Framing

##

-   **Framing** is a concept initially coined by [@bateson_steps_2000] and developed by [@goffman_frame_1986].

    -   It is defined as the interpretation schemes that individuals use to understand reality and act based on the interpretation they perform.

-   In general terms, this means that individuals have built mental filters throughout their lives that allow them to understand the world. In turn, these mental filters influence the decisions they make.

## {.smaller}

-   Example of how **framing** can change decisions depending on how information is presented [@tversky_framing_1981, p 453]:

    -   **Problem 1** \[N = 152\][^1]: Imagine that the U.S. is preparing for the outbreak of an unusual Asian disease, which is expected to kill 600 people. Two alternative programs to combat the disease have been proposed. Assume that the exact scientific estimate of the consequences of the programs are as follows:

        -   If Program A is adopted, 200 people will be saved. **\[72 percent\]**[^2]
        -   If Program B is adopted, there is 1/3 probability that 600 people will be saved, and 2/3 probability that no people will be saved. **\[28 percent\]**[^3]

    -   Which of the two programs would you favor?

[^1]: Number of people that participate in the experiment

[^2]: Percentage of people that choose Program A

[^3]: Percentage of people that choose Program B

## {.smaller}

-   Example of how **framing** can change decisions depending on how information is presented [@tversky_framing_1981, p 453]:

    -   **Problem 2** \[N = 155\][^4]: The same statement as **Problem 1**

        -   If Program C is adopted, 400 people will die. **\[22 percent\]**[^5]
        -   If Program D is adopted there is 1/3 probability that nobody will die, and 2/3 probability that 600 people will die. **\[78 percent\]**[^6]

    -   Which of the two programs would you favor?

-   **Problem 1** and **Problem 2** are equivalent but they are frame in a different way. This is why the majority choice in **Problem 1** is Program A and for the **Problem 2** is Program D.

[^4]: Number of people that participate in the experiment

[^5]: Percentage of people that choose Program C

[^6]: Percentage of people that choose Program D

##

-   In the field of negotiation you can also influence the perception of the counterpart through framing. An introduction to this topic can be found in [@shonk_framing_2020]:

    -   Offer Manageable Choices [@iyengar_when_2000]
    -   Make Several Offers [@leonardelli_multiple_2019]
    -   Be Willing to Be Rejected [@simonson_choice_1992]

# Cognitive Biases in Negotiation

## {.smaller}

-   A **cognitive bias** is defined as a systematic error that affects the decisions and judgments.

```{mermaid}
%%| label: fig-cognitive-biases-in-negotiation-1
%%| fig-cap: Cognitive biases in the context of a negotiation[^7]

flowchart TB
    A("Cognitive biases in negotiation") --> A1("Irrational Escalation of Commitment")
    A --> A2("Mythical fixed-pie beliefs")
    A --> A3("Anchoring and Adjustment")
    A --> A4("Issue Framing and Risk")
    A --> A5("Availability of Information")
    A --> A6("The Winner’s Curse")
    A1 --> A11("Persist in pursuing a failing course of action")
    A2 --> A21("Tendency to see negotiation as a zero-sum situation")
    A3 --> A31("Be influenced by a reference point and failing to make adjustements")
    A4 --> A41("Be influenced by a frame in a risk situation")
    A5 --> A51("Tendency to overweight information that is easily recalled")
    A6 --> A61("Feel discomfort about a negotiation win")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A2,A21,A3,A31,A4,A41,A5,A51,A6,A61 UmngNode;
```

[^7]: For a general **cognitive bias** taxonomy check out [@dimara_task-based_2020]

## {.smaller}

-   A **cognitive bias** is defined as a systematic error that affects the decisions and judgments.

```{mermaid}
%%| label: fig-cognitive-biases-in-negotiation-2
%%| fig-cap: Cognitive biases in the context of a negotiation[^8]

flowchart TB
    A("Cognitive biases in negotiation") --> A1("Overconfidence")
    A --> A2("The Law of Small Numbers")
    A --> A3("Self-Serving Biases")
    A --> A4("Endowment Effect")
    A --> A5("Ignoring others’ cognitions")
    A --> A6("Reactive devaluation")
    A1 --> A11("Tendency to believe that one’s ability to be accurate is greater than it is")
    A2 --> A21("Draw conclusions based on small data samples")
    A3 --> A31("Overestimate the internal factors and underestimate the external factors")
    A4 --> A41("Tendency to inflate the value of something you own")
    A5 --> A51("Failure to consider the other party’s")
    A6 --> A61("Placing less value on concessions because the other party offered them")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A2,A21,A3,A31,A4,A41,A5,A51,A6,A61 UmngNode;
```

[^8]: For a general **cognitive bias** taxonomy check out [@dimara_task-based_2020]

# An alternative approach: Mood and Emotion

##

-   The study of **cognitive biases** in negotiation is focused on how negotiators make judgment errors or how can negotiators can mitigate or eliminate this errors to improve the decision making process.

-   Another approach is to use emotions as a strategic tool in negotiations[^9] by adjusting the negotiation stance based on the other party’s emotional state.

[^9]: As a conflict negotiation professor I am quite bad at using this tool

##

```{mermaid}
%%| label: fig-positive-emotions-consequences
%%| fig-cap: Positive emotions and consequences for negotiations [@lewicki_negociacion_2024, p. 202]

flowchart LR
    A("Why") --> A1("Positive emotions")
    A1 --> A11("Generate positive consequences for negotiations")
    A11 --> A111("Lead the parties toward more integrative processes")
    A11 --> A112("Create a positive attitude toward the other side")
    A11 --> A113("Promote persistence")
    A11 --> A114("Set the stage for successful subsequent negotiations")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A111,A112,A113,A114 UmngNode;
```

##

```{mermaid}
%%| label: fig-aspects-negotiation-process-positive-emotions
%%| fig-cap: Aspects of the negotiation process and positive emotions [@lewicki_negociacion_2024, p. 203]
%%| fig-height: 4

flowchart LR
    A("Why") --> A1("Aspects of the Negotiation Process")
    A1 --> A11("Fair procedures during negotiation")
    A1 --> A12("Favorable social comparisons (Comparing your outcome to others outside the negotiation)")
    A11 --> A111("Positive emotions")
    A12 --> A111

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A111,A12 UmngNode;
```

##

```{mermaid}
%%| label: fig-negative-emotions-consequences
%%| fig-cap: Negative emotions and consequences for negotiations [@lewicki_negociacion_2024, pp. 203-204]
%%| fig-height: 5

flowchart LR
    A("Why") --> A1("Negative emotions")
    A1 --> A11("Generate negative consequences for negotiations")
    A11 --> A111("Lead the parties to define the situation as distributive")
    A11 --> A112("Undermine a negotiator’s ability to analyze the situation accurately affecting individual outcome")
    A11 --> A113("Lead the parties to escalate the conflict")
    A11 --> A114("Lead the parties to retaliate and prevent integrative outcomes")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A111,A112,A113,A114 UmngNode;
```

##

```{mermaid}
%%| label: fig-aspects-negotiation-process-negative-emotions
%%| fig-cap: Aspects of the negotiation process and negative emotions [@lewicki_negociacion_2024, p. 205]

flowchart LR
    A("Why") --> A1("Aspects of the Negotiation Process")
    A1 --> A11("Competitive mindset")
    A1 --> A12("Negotiation ending in an impasse")
    A1 --> A13("Negative anxiety from the prospect of beginning a negotiation")
    A1 --> A14("Negative experiences carry over from past negotiations")
    A11 --> A111("Negative emotions")
    A12 --> A111
    A13 --> A111
    A14 --> A111

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A111,A12,A13,A14 UmngNode;
```

# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html) and [**$\pi$-thon**](https://www.python.org/about/)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://rstudio.com/products/rstudio/), [**Quarto**](https://quarto.org/) and the authors and maintainers of [**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

# References {.allowframebreaks}