---
title: Finding and Using Negotiation Power
subtitle: Session 1
author:
  - name: 
      given: Luis F.
      family: Gómez L.
    orcid: 0000-0002-2877-9015
    email: luis.gomezl@unimilitar.edu.co
    affiliations:
      - name: Distance Learning Faculty
toc: true
toc-depth: 1
slide-level: 2
format:
  presentation-umng-revealjs:
    logo-vertical-strip: 000_logos/logo-white-vertical.png 
    embed-resources: true
    mermaid-format: png
    reference-location: document
date: last-modified
date-format: "D MMMM, YYYY"
knitr: 
  opts_chunk: 
    echo: false
    warning: false
    fig-align: center
    out-width: 85%
lang: en
bibliography: conflict_neg.bib
csl: apa-7.csl
editor: source
---

```{r}
#| label: libraries


```

```{r}
#| label: palette

umng_palette <- c(
  "#043074",
  "#fdc600",
  "#0d5e30",
  "#ee2a24",
  "#fc6700",
  "#00b3f0",
  "#6e3c1a",
  "#f8941c",
  "#8e44ad",
  "#2c3e50",
  "#16a085",
  "#c0392b",
  "#e91e63"
)
```

# Please Read Me

##

-   Check the message **Welcome greeting** published in the News Bulletin Board.

-   Dear student please edit your profile uploading a photo where your face is clearly visible.

-   The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

-   This presentation is based on [@lewicki_negociacion_2024, Chapter 8]

##

- **Purpose**

    - Understand the role of power and the different sources from which this element arises in a negotiation.

# Power in the context of a negotiation

##

-   Power in the context of a negotiation refers to the ability of a negotiator to gain an advantage or increase the likelihood of approaching its target point.

-   Power is important in a negotiation because it generates advantages and allows reaching a settlement point close to the target point.

##

```{mermaid}
%%| label: fig-perspectives-about-power
%%| fig-cap: Perspectives about power[^1]
%%| fig-height: 4

flowchart LR
    A("2 perspectives") --> A1("Power over")
    A --> A2("Power with")
    A1 --> A11("Power is fundamentally dominating and coercive in nature")
    A2 --> A21("Power holder jointly develops and shares power with the other")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A2,A21 UmngNode;
```

[^1]: Check out [@coleman_power_2014] if you want a general perspective about power and its relation with conflict

# Sources of power in a negotiation

## 

```{mermaid}
%%| label: fig-sources-of-power-1
%%| fig-cap: Sources of power in a negotiation [@lewicki_negociacion_2024, pp. 242-260]
%%| fig-height: 4

flowchart TB
    A("Sources") --> A1("Informational")
    A --> A2("Based on Personality and Individual Differences")
    A2 --> A21("Cognitive Orientation")
    A2 --> A22("Motivational Orientation")
    A2 --> A23("Dispositions and Related Skills")
    A2 --> A24("Moral Orientation toward Power")
    A2 --> A25("Moods")
    A --> A3("Structural Power")
    A3 --> A31("Hierarchy")
    A31 --> A311("Legitimate power")
    A31 --> A312("Resource Power")
    A3 --> A32("Location in a Network")
    A32 --> A321("Field: Social network analysis")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A2,A21,A22,A23,A24,A25,A3,A31,A311,A312,A32,A321 UmngNode;
```

## {.center}

```{mermaid}
%%| label: fig-sources-of-power-2
%%| fig-cap: Sources of power in a negotiation [@lewicki_negociacion_2024, pp. 242-260]

flowchart TB
    A("Sources") --> A1("Based on Relationships")
    A1 --> A11("Goal Interdependence")
    A1 --> A12("Referent Power")
    A --> A2("Contextual")
    A2 --> A21("BATNAs")
    A2 --> A22("Culture")
    A2 --> A23("Agents, Constituencies and External Audiences")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A12,A2,A21,A22,A23 UmngNode;
```

# The Consequences of Unequal Power

##

```{mermaid}
%%| label: fig-consequences-of-unequal-power
%%| fig-cap: Consequences of unequal power in a negotiation [@lewicki_negociacion_2024, pp. 260-261]

flowchart TB
    A("Unequal power") --> A1("Consequences")
    A1 --> A11("Lead to significantly different conflict orientations and behaviors")
    A11 --> A111("Avoid making things worse")
    A11 --> A112("Exit the relationship")
    A11 --> A113("Hinder the other's actions")
    A1 --> A12("More likely to behave contentiously")
    A12 --> A121("Threats")
    A12 --> A122("Punishments")
    A12 --> A123("Exploitative behavior")
    A1 --> A13("The more powerful party has the capacity to determine the negotiation outcome")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A111,A112,A113,A12,A121,A122,A123,A13 UmngNode;    
```

# Dealing with negotatiors how have more power

## {.smaller}

-   Advice to negotiators who are in low-power positions based on [@lewicki_negociacion_2024, pp. 261-262]:

    -   Diversify risk by entering into deals with several other partners.
    -   Deal with a variety of different individuals and departments in the high-power party (Divide and Conquer).
    -   Build coalitions with other low-power players to increase collective bargaining power.
    -   Enter in early deals with high-power parties and maximize the visibility of those deals to other parties.

## {.smaller}

-   Advice to negotiators who are in low-power positions based on [@lewicki_negociacion_2024, pp. 261-262]:

    -   If you have something to offer, make sure you offer it to more than one high-power party to generate competition between them.
    -   Gather and leverage relevant information to strengthen your negotiation position and achieve better outcomes through persuasive communication.
    -   Do what you can to manage the process (for example the agenda or location) to guide the negotiation towards a more favorable outcome.

# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html) and [**$\pi$-thon**](https://www.python.org/about/)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://rstudio.com/products/rstudio/), [**Quarto**](https://quarto.org/) and the authors and maintainers of the packages
[**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

# References {.allowframebreaks}