---
title: "Finding and Using Negotiation Power"
author: "Luis Francisco Gomez Lopez"
date: 2022-06-13 10:29:10
output:
  beamer_presentation:
    colortheme: dolphin
    fonttheme: structurebold
    theme: AnnArbor
    fig_caption: true
  slidy_presentation: default
bibliography: conflict_neg.bib
link-citations: yes
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{fontawesome5}
  - \usepackage{xcolor}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      message   = FALSE,
                      fig.align = "center")
```

```{r libraries}

library(tidyverse)
library(DiagrammeR)
```

# Contents

- Please Read Me
- Purpose
- Power in the context of a negotiation
- Sources of power in a negotiation
- Dealing with negotiators how have more power
- Acknowledgments
- References

# Please Read Me

- Check the message __Welcome greeting__ published in the News Bulletin Board.

- Dear student please edit your profile uploading a photo where your face is clearly visible.

- The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

- This presentation is based on [@lewicki_essentials_2016, Chapter 8]

# Purpose

Understand the role of power and the different sources from which this element arises in a negotiation.

# Power in the context of a negotiation

- Power in the context of a negotiation refers to the ability of a negotiator to gain an advantage or increase the likelihood of approaching its target point.

- Power is important in a negotiation because it generates advantages and allows reaching a settlement point close to the target point.

# Power in the context of a negotiation

- Perspectives about power[^1]

```{r fig_1, out.width = '85%'}

grViz(diagram = '
      
      digraph boxes_and_circles {
      
      graph [rankdir = LR]
      
      node[shape = circle, color = "#2C3E50"]
      
      "2\nperspectives";
      
      node[shape = oval, color = "#E31A1C"]
      
      "Power\nover";
      "Power\nwith";
      
      node[shape = oval, color = "#18BC9C"]
      
      "Power is fundamentally\ndominating and coercive\nin nature";
      "Power holder jointly\ndevelops and shares power\nwith the other";
      
      "2\nperspectives" -> "Power\nover";
      "2\nperspectives" -> "Power\nwith";
      
      "Power\nover" -> "Power is fundamentally\ndominating and coercive\nin nature";
      "Power\nwith" -> "Power holder jointly\ndevelops and shares power\nwith the other"
      
      }
      
      ')
```

[^1]: Check out [@coleman_power_2014] if you want a general perspective about power and its relation with conflict 

# Sources of power in a negotiation

```{r fig_2, out.width = '80%'}

grViz(diagram = '
      
      digraph boxes_and_circles {
      
      graph [rankdir = LR]
      
      node[shape = circle, color = "#2C3E50"]
      
      Sources;
      
      node[shape = oval, color = "#E31A1C"]
      
      Informational;
      "Based on\nPersonality and\nIndividual Differences";
      "Structural\nPower";
      
      node[shape = oval, color = "#18BC9C"]
      
      "Cognitive\nOrientation";
      "Motivational\nOrientation";
      "Dispositions\nand\nRelated Skills";
      "Moral Orientation\ntoward Power";
      Moods;
      
      Hierarchy;
      "Location in\na Network";
      
      node[shape = oval, color = "#CCBE93"]
      
      "Legitimate\npower";
      "Resource\nPower";
      "Field:\nSocial network\nanalysis";
      
      Sources -> Informational;
      Sources -> "Based on\nPersonality and\nIndividual Differences";
      Sources -> "Structural\nPower";
      
      "Based on\nPersonality and\nIndividual Differences" -> "Cognitive\nOrientation";
      "Based on\nPersonality and\nIndividual Differences" -> "Motivational\nOrientation";
      "Based on\nPersonality and\nIndividual Differences" -> "Dispositions\nand\nRelated Skills";
      "Based on\nPersonality and\nIndividual Differences" -> "Moral Orientation\ntoward Power";
      "Based on\nPersonality and\nIndividual Differences" -> Moods;
      "Structural\nPower" -> Hierarchy;
      "Structural\nPower" -> "Location in\na Network";
      
      Hierarchy -> "Legitimate\npower"
      Hierarchy -> "Resource\nPower"
      "Location in\na Network" -> "Field:\nSocial network\nanalysis"
      
      }
      
      ')
```

# Sources of power in a negotiation

```{r fig_3, out.width = '75%'}

grViz(diagram = '
      
      digraph boxes_and_circles {
      
      graph [rankdir = LR]
      
      node[shape = circle, oval = "#2C3E50"]
      
      Sources;
      
      node[shape = oval, color = "#E31A1C"]
      
      "Based on\nRelationships";
      Contextual;
      
      node[shape = oval, color = "#18BC9C"]
      
      "Goal\nInterdependence";
      "Referent\nPower";
      
      BATNAs;
      Culture;
      "Agents,\nConstituencies\nand\nExternal Audiences";
      
      Sources -> "Based on\nRelationships";
      Sources -> Contextual

      "Based on\nRelationships" -> "Goal\nInterdependence";
      "Based on\nRelationships" -> "Referent\nPower";
      Contextual -> BATNAs;
      Contextual -> Culture;
      Contextual -> "Agents,\nConstituencies\nand\nExternal Audiences";
      
      }
      
      ')
```

# Dealing with negotatiors how have more power 

- Advice to negotiators who are in low-power positions based on [@lewicki_essentials_2016, Chapter 8, p 199-201]:

    + Diversify risk by entering into deals with several other partners.
    + Deal with a variety of different individuals and departments in the high-power party (Divide and Conquer).
    + Build coalitions with other low-power players to increase collective bargaining power.
    + If you have something to offer, make sure you offer it to more than one high-power party to generate competition between them.

# Acknowledgments

- To my family that supports me

- To the taxpayers of Colombia and the __[UMNG students](https://www.umng.edu.co/estudiante)__ who pay my salary

- To the __[Business Science](https://www.business-science.io/)__ and __[R4DS Online Learning](https://www.rfordatasci.com/)__ communities where I learn __[R](https://www.r-project.org/about.html)__ 

- To the __[R Core Team](https://www.r-project.org/contributors.html)__, the creators of __[RStudio IDE](https://rstudio.com/products/rstudio/)__ and the authors and maintainers of the packages __[tidyverse](https://CRAN.R-project.org/package=tidyverse)__, __[DiagrammeR](https://CRAN.R-project.org/package=DiagrammeR)__, __[knitr](https://CRAN.R-project.org/package=knitr)__, __[kableExtra](https://CRAN.R-project.org/package=kableExtra)__ and __[tinytex](https://CRAN.R-project.org/package=tinytex)__ for allowing me to access these tools without paying for a license

- To the __[Linux kernel community](https://www.kernel.org/category/about.html)__ for allowing me the possibility to use some __[Linux distributions](https://static.lwn.net/Distributions/)__ as my main __[OS](https://en.wikipedia.org/wiki/Operating_system)__ without paying for a license

# References
