---
title: "Ethics in Negotiation"
author: "Luis Francisco Gomez Lopez"
date: 2022-06-13 10:30:50
output:
  beamer_presentation:
    colortheme: dolphin
    fonttheme: structurebold
    theme: AnnArbor
    fig_caption: true
  slidy_presentation: default
bibliography: conflict_neg.bib
link-citations: yes
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{fontawesome5}
  - \usepackage{xcolor}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      message   = FALSE,
                      fig.align = "center")
```

```{r libraries}

library(tidyverse)
library(DiagrammeR)
```

# Contents

- Please Read Me
- Purpose
- Ethics and its relationship with negotiation
- Ethically ambiguous tactics
- Motives and consequences of using deceptive tactics
- Dealing with the use of deceptive tactics by the counterpart
- Acknowledgments
- References

# Please Read Me

- Check the message __Welcome greeting__ published in the News Bulletin Board.

- Dear student please edit your profile uploading a photo where your face is clearly visible.

- The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

- This presentation is based on [@lewicki_essentials_2016, Chapter 5]

# Purpose

Explore and understand the ethical standards commonly accepted in a negotiation process in order to detect and deal with deceptive tactics.

# Ethics and its relationship with negotiation

- Ethics is understand as the social standards that are apply to examine what is right or wrong in a specific situation or a process to establish such standards [@lewicki_essentials_2016, Chapter 5, p 116].

- The ethical considerations in a negotiation are related to how the exchange of information occurs [@lewicki_ethical_1998].

    + Because the exchange of information in the negotiation process is vital, the analysis of ethics is associated with examining whether or not there is a dishonest communication.
    
# Ethics and its relationship with negotiation

- To evaluate how ethical the strategies and tactics are in a negotiation, 4 standards can be used [@lewicki_essentials_2016, Chapter 5, p 118]:

```{r fig_1, out.width = '90%'}

grViz(diagram = '
      
      digraph boxes_and_circles {
      
      graph [rankdir = TB]
      
      node[shape = circle, color = "#2C3E50"]
      
      "Ethical standards";
      
      node[shape = oval, color = "#E31A1C"]
      
      "End-result ethics";
      "Duty ethics";
      "Social contract ethics";
      "Personalistic ethics";
      
      node[shape = oval, color = "#18BC9C"]
      
      "Jeremy Bentham\nJohn Stuart Mill";
      "Immanuel Kant";
      "Jean-Jacques Rousseau";
      "Martin Buber"
      
      "Ethical standards" -> "End-result ethics";
      "Ethical standards" -> "Duty ethics";
      "Ethical standards" -> "Social contract ethics";
      "Ethical standards" -> "Personalistic ethics";
      
       "End-result ethics" -> "Jeremy Bentham\nJohn Stuart Mill";
       "Duty ethics" -> "Immanuel Kant";
       "Social contract ethics" -> "Jean-Jacques Rousseau";
       "Personalistic ethics" -> "Martin Buber"
      
      }
      
      ')
```

# Ethically ambiguous tactics

- There are tactics that are not ethical and that can be quickly identified, such as stealing confidential data from the counterpart. 

    + These types of tactics must be removed from the negotiator's toolbox.

- However, in the context of negotiation there are gray areas. These gray areas are known as ethically ambiguous tactics within the theory of negotiation. 

    + These tactics are related to what the negotiators say or what they claim they will do concerning what they really do.
    
# Ethically ambiguous tactics

- Gray areas within a negotiating context regarding ethically ambiguous tactics are presented due to the 2 dilemmas a negotiator faces:

    + Dilemma of honesty
    
        + How much truth should be revealed to the counterpart?
    
    + Dilemma of trust
    
        + How much should a negotiator believe what the counterpart says?

# Ethically ambiguous tactics

- Types of ethically ambiguous tactics

```{r fig_2, out.width = '100%'}

grViz(diagram = '
      
      digraph boxes_and_circles {
      
      graph [rankdir = LR]
      
      node[shape = circle, color = "#2C3E50"]
      
      "Dimensions used\nin the assessment of\nethically ambiguous tactics";
      
      node[shape = oval, color = ForestGreen]
      
      "Appropriateness";
      "Likelihood\nof using it";
      "Perceived efficacy\nof using it";
      
      node[shape = circle, color = "#E31A1C"]
      
      "Types"
      
      node[shape = oval, color = "#18BC9C"]
      
      "Traditional competitive\nbargaining";
      "Emotional\nmanipulation";
      "Misrepresentation";
      "Misrepresentation to\nopponent’s networks";
      "Inappropriate information\ngathering";
      "Bluffing";
      
      node[shape = oval, color = "#CCBE93"]
      
      "Perceived as\nappropriate and\nlikely to be used"
      
      "Dimensions used\nin the assessment of\nethically ambiguous tactics" -> "Appropriateness";
      "Dimensions used\nin the assessment of\nethically ambiguous tactics" -> "Likelihood\nof using it";
      "Dimensions used\nin the assessment of\nethically ambiguous tactics" -> "Perceived efficacy\nof using it";
      
      "Appropriateness" -> "Types";
      "Likelihood\nof using it" -> "Types";
      "Perceived efficacy\nof using it" -> "Types";

      "Types" -> "Traditional competitive\nbargaining";
      "Types" -> "Emotional\nmanipulation";
      "Types" -> "Misrepresentation";
      "Types" -> "Misrepresentation to\nopponent’s networks";
      "Types" -> "Inappropriate information\ngathering";
      "Types" -> "Bluffing";

      "Traditional competitive\nbargaining" -> "Perceived as\nappropriate and\nlikely to be used";
      "Emotional\nmanipulation" -> "Perceived as\nappropriate and\nlikely to be used";
      
      }
      
      ')
```

# Motives and consequences of using deceptive tactics

```{r fig_3, out.width = '60%'}

grViz(diagram = '
      
      digraph boxes_and_circles {
      
      graph [rankdir = TB]
      
      node[shape = circle, color = "#2C3E50"]
      
      "Deceptive\ntactics";
      
      node[shape = oval, color = "#E31A1C"]
      
      "Types"
      
      node[shape = oval, color = "#18BC9C"]
      
      "Active";
      "Pasive";
      
      node[shape = oval, color = "#CCBE93"]
      
      "Misrepresent\nlying openly";
      "Misrepresent by omission\nnot revealing information\nthat benefits the\ncounterpart";

      "Deceptive\ntactics" -> "Types";
      
      "Types" -> "Active";
      "Types" -> "Pasive";
      
      "Active" -> "Misrepresent\nlying openly";
      "Pasive" -> "Misrepresent by omission\nnot revealing information\nthat benefits the\ncounterpart";

      }
      
      ')
```

# Motives and consequences of using deceptive tactics

- Why do negotiators use deceptive tactics in the context of a negotiation?

    + Need to acquire greater power through the manipulation of information to get closer to the target point
    
    + Use of a more competitive negotiation style[^1]
    
[^1]: This aspect generates a greater probability of using this type of tactics

# Motives and consequences of using deceptive tactics

```{r fig_4, out.width = '65%'}

grViz(diagram = '
      
      digraph boxes_and_circles {
      
      graph [rankdir = TB]
      
      node[shape = circle, color = "#2C3E50"]
      
      "Consequences of\nUnethical Conduct";
      
      node[shape = oval, color = "#E31A1C"]
      
      "Positive\nor Negative";
      
      node[shape = oval, color = "#18BC9C"]
      
      "Aspects";
      
      node[shape = oval, color = "#CCBE93"]
      
      "Effectiveness";
      "Reactions of Others";
      "Reactions of Self"

      "Consequences of\nUnethical Conduct" -> "Positive\nor Negative";
      
      "Positive\nor Negative" -> "Aspects";
      
      "Aspects" -> "Effectiveness";
      "Aspects" -> "Reactions of Others";
      "Aspects" -> "Reactions of Self"

      }
      
      ')
```

# Motives and consequences of using deceptive tactics

- **Effectiveness**

    + Deceptive tactics generate **positive consequences** when the outcome of the negotiation improves compared to whether a negotiator had acted ethically and if the conduct is not punished where the consequences materialize in the short run.
    
    + Deceptive tactics generate **negative consequences** because the reputation of the negotiator is damaged where the consequences materialize in the future.
    
# Motives and consequences of using deceptive tactics

- **Reactions of Others**

    + Deceptive tactics generate **positive consequences** only if constituents, indirect actors or interest observers considered appropriate to use this type of tactics[^2].
    
    + Deceptive tactics generate **negative consequences** because retaliations occur directly from the counterpart and possibly from the constituents, indirect actors or interest observers if they consider that the tactic used is inappropriate[^3]. 
    
[^2]: This is my own personal opinion and it is not mentioned in [@lewicki_essentials_2016, Chapter 5]
[^3]: As a personal opinion in the case of the indirect actors or interest observers the retaliation is materialize through a social sanction.

# Motives and consequences of using deceptive tactics

- **Reactions of Self**

    + Deceptive tactics generate **positive consequences** only if the negotiator does not suffer from guilt, remorse or discomfort.
    
    + Deceptive tactics generate **negative consequences** when the negotiator suffers from guilt, remorse or discomfort for having used these tactics[^4]

[^4]: This issue directly affects the negotiation since the negotiator is willing to make greater concessions to the counterpart to compensate for using deceptive tactics.

# Dealing with the use of deceptive tactics by the counterpart

```{r fig_5, out.width = '45%'}

grViz(diagram = '
      
      digraph boxes_and_circles {
      
      graph [rankdir = LR]
      
      node[shape = circle, color = "#2C3E50"]
      
      Options;
      
      node[shape = oval, color = "#E31A1C"]
      
      "Ask Probing Questions";
      "Phrase Questions\nin Different Ways";
      "Force the Other Party to\nLie or Back Off";
      "Test the Other Party";
      "“Call” the Tactic";
      "Ignore the Tactic";
      "Discuss What You See and Offer to\nHelp the Other Party Shift\nto More Honest Behaviors";
      "Respond in Kind"
      
      Options -> "Ask Probing Questions";
      Options -> "Phrase Questions\nin Different Ways";
      Options -> "Force the Other Party to\nLie or Back Off";
      Options -> "Test the Other Party";
      Options -> "“Call” the Tactic";
      Options -> "Ignore the Tactic";
      Options -> "Discuss What You See and Offer to\nHelp the Other Party Shift\nto More Honest Behaviors";
      Options -> "Respond in Kind"

      }
      
      ')
```

# Dealing with the use of deceptive tactics by the counterpart

- **My personal opinion**:

    + Read the article [@adler_negotiating_2007]:
    
        + Before the Bargaining Begins
        
        + During the Bargaining Process
        
    + I think the best approach is to follow the recommendations in the section **Before the Bargaining Begins** because they are carried out during the planning stage of a negotiation where there is more time and information that can be collected in order to respond adequately to deceptive tactics.

# Acknowledgments

- To my family that supports me

- To the taxpayers of Colombia and the __[UMNG students](https://www.umng.edu.co/estudiante)__ who pay my salary

- To the __[Business Science](https://www.business-science.io/)__ and __[R4DS Online Learning](https://www.rfordatasci.com/)__ communities where I learn __[R](https://www.r-project.org/about.html)__ 

- To the __[R Core Team](https://www.r-project.org/contributors.html)__, the creators of __[RStudio IDE](https://rstudio.com/products/rstudio/)__ and the authors and maintainers of the packages __[tidyverse](https://CRAN.R-project.org/package=tidyverse)__, __[DiagrammeR](https://CRAN.R-project.org/package=DiagrammeR)__ and __[tinytex](https://CRAN.R-project.org/package=tinytex)__ for allowing me to access these tools without paying for a license

- To the __[Linux kernel community](https://www.kernel.org/category/about.html)__ for allowing me the possibility to use some __[Linux distributions](https://static.lwn.net/Distributions/)__ as my main __[OS](https://en.wikipedia.org/wiki/Operating_system)__ without paying for a license

# References
 