---
title: Ethics in Negotiation
subtitle: Session 1
author:
  - name: 
      given: Luis F.
      family: Gómez L.
    orcid: 0000-0002-2877-9015
    email: luis.gomezl@unimilitar.edu.co
    affiliations:
      - name: Distance Learning Faculty
toc: true
toc-depth: 1
slide-level: 2
format:
  presentation-umng-revealjs:
    logo-vertical-strip: 000_logos/logo-white-vertical.png 
    embed-resources: true
    mermaid-format: png
    reference-location: document
date: last-modified
date-format: "D MMMM, YYYY"
knitr: 
  opts_chunk: 
    echo: false
    warning: false
    fig-align: center
    out-width: 85%
lang: en
bibliography: conflict_neg.bib
csl: apa-7.csl
editor: source
---

```{r}
#| label: libraries


```

```{r}
#| label: palette

umng_palette <- c(
  "#043074",
  "#fdc600",
  "#0d5e30",
  "#ee2a24",
  "#fc6700",
  "#00b3f0",
  "#6e3c1a",
  "#f8941c",
  "#8e44ad",
  "#2c3e50",
  "#16a085",
  "#c0392b",
  "#e91e63"
)
```

# Please Read Me

##

-   Check the message **Welcome greeting** published in the News Bulletin Board.

-   Dear student please edit your profile uploading a photo where your face is clearly visible.

-   The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

-   This presentation is based on [@lewicki_negociacion_2024, Chapter 5]

##

- **Purpose**

    - Explore and understand the ethical standards commonly accepted in a negotiation process in order to detect and deal with deceptive tactics.

# Ethics and its relationship with negotiation

##

-   Ethics is understand as the social standards that are apply to examine what is right or wrong in a specific situation or a process to establish such standards [@lewicki_negociacion_2024, Chapter 5, p. 136].

-   The ethical considerations in a negotiation are related to how the exchange of information occurs [@lewicki_ethical_1998].

    -   Because the exchange of information in the negotiation process is vital, the analysis of ethics is associated with examining whether or not there is a dishonest communication.

##

-   To evaluate how ethical the strategies and tactics are in a negotiation, 4 standards can be used [@lewicki_negociacion_2024, Chapter 5, p. 118]:

```{mermaid}
%%| label: fig-approaches-to-ethical-reasoning
%%| fig-cap: Approaches to ethical reasoning to evaluate strategies and tactics in a negotiation [@lewicki_negociacion_2024, pp. 140]

flowchart TB
    A("Ethical standards") --> A1("End-result ethics")
    A1 --> A11("`Jeremy Bentham 
    John Stuart Mill`")
    A --> A2("Duty ethics")
    A2 --> A21("Immanuel Kant")
    A --> A3("Social contract ethics")
    A3 --> A31("Jean-Jacques Rousseau")
    A --> A4("Personalistic ethics")
    A4 --> A41("Martin Buber")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A2,A21,A3,A31,A4,A41 UmngNode;
```

# Ethically ambiguous tactics

##

-   There are tactics that are not ethical and that can be quickly identified, such as stealing confidential data from the counterpart.

    -   These types of tactics must be removed from the negotiator's toolbox.

-   However, in the context of negotiation there are gray areas. These gray areas are known as ethically ambiguous tactics within the theory of negotiation.

    -   These tactics are related to what the negotiators say or what they claim they will do concerning what they really do.
    
##

-   Gray areas within a negotiating context regarding ethically ambiguous tactics are presented due to the 2 dilemmas a negotiator faces:

    -   Dilemma of honesty

        -   How much truth should be revealed to the counterpart?

    -   Dilemma of trust

        -   How much should a negotiator believe what the counterpart says?

##

```{mermaid}
%%| label: fig-types-ethically-ambiguous-tactics
%%| fig-cap: Types of ethically ambiguous tactics [@lewicki_negociacion_2024, p. 148]
%%| fig-height: 5.5

flowchart TB
    A("Dimensions used in the assessment of ethically ambiguous tactics") --> A1("Appropriateness")
    A1 --> A11("Types")
    A11 --> A111("Traditional competitive bargaining")
    A111 --> A1111("Perceived as appropriate and likely to be used")
    A11 --> A112("Emotional manipulation")
    A112 --> A1111
    A11 --> A113("Misrepresentation")
    A11 --> A114("Misrepresentation to opponent’s networks")
    A11 --> A115("Inappropriate information gathering")
    A11 --> A116("Bluffing")
    A --> A2("Likelihood of using it")
    A2 --> A11
    A --> A3("Perceived efficacy of using it")
    A3 --> A11

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A111,A1111,A112,A113,A114,A115,A116,A2,A3 UmngNode;
```

# Motives and consequences of using deceptive tactics

##

```{mermaid}
%%| label: fig-use-of-deceptive-tactics
%%| fig-cap: Use of deceptive tactics [@lewicki_negociacion_2024, p. 151]

flowchart TB
    A("Deceptive tactics") --> A1("Types")
    A1 --> A11("Active")
    A11 --> A111("Misrepresent lying openly")
    A1 --> A12("Pasive")
    A12 --> A121("Misrepresent by omission not revealing information that benefits the counterpart")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A111,A12,A121 UmngNode;
```

##

-   Why do negotiators use deceptive tactics in the context of a negotiation?

    -   Need to acquire greater power through the manipulation of information to get closer to the target point

    -   Use of a more competitive negotiation style[^1]

[^1]: This aspect generates a greater probability of using this type of tactics

##

```{mermaid}
%%| label: fig-consequences-of-unethical-conduct
%%| fig-cap: Consequences of unethical conduct [@lewicki_negociacion_2024, pp. 155-158]

flowchart TB
    A("Consequences of Unethical Conduct") --> A1("Types")
    A1 --> A11("Positive or Negative")
    A11 --> A111("Aspects")
    A111 --> A1111("Effectiveness")
    A111 --> A1112("Reactions of Others")
    A111 --> A1113("Reactions of Self")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A11,A111,A1111,A1112,A1113 UmngNode;
```

##

-   **Effectiveness**

    -   Deceptive tactics generate **positive consequences** when the outcome of the negotiation improves compared to whether a negotiator had acted ethically and if the conduct is not punished where the consequences materialize in the short run.

    -   Deceptive tactics generate **negative consequences** because the reputation of the negotiator is damaged where the consequences materialize in the future.

##

-   **Reactions of Others**

    -   Deceptive tactics generate **positive consequences** only if constituents, indirect actors or interest observers considered appropriate to use this type of tactics[^2].

    -   Deceptive tactics generate **negative consequences** because retaliations occur directly from the counterpart and possibly from the constituents, indirect actors or interest observers if they consider that the tactic used is inappropriate[^3].

[^2]: This is my own personal opinion and it is not mentioned in [@lewicki_negociacion_2024, Chapter 5]

[^3]: As a personal opinion in the case of the indirect actors or interest observers the retaliation is materialize through a social sanction.

##

-   **Reactions of Self**

    -   Deceptive tactics generate **positive consequences** only if the negotiator does not suffer from guilt, remorse or discomfort.

    -   Deceptive tactics generate **negative consequences** when the negotiator suffers from guilt, remorse or discomfort for having used these tactics[^4]

[^4]: This issue directly affects the negotiation since the negotiator is willing to make greater concessions to the counterpart to compensate for using deceptive tactics.

# Dealing with the use of deceptive tactics by the counterpart

## {.center}

```{mermaid}
%%| label: fig-dealing-with-deceptive-tactics-1
%%| fig-cap: Dealing with deceptive tactics by the counterpart [@lewicki_negociacion_2024, pp. 172-174]

flowchart TB
    A("Options") --> A1("Ask Probing Questions")
    A --> A2("Phrase Questions in Different Ways")
    A --> A3("Force the Other Party to Lie or Back Off")
    A --> A4("Test the Other Party")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A2,A3,A4 UmngNode;
```

## {.center}

```{mermaid}
%%| label: fig-dealing-with-deceptive-tactics-2
%%| fig-cap: Dealing with deceptive tactics by the counterpart [@lewicki_negociacion_2024, pp. 172-174]

flowchart TB
    A("Options") --> A1("“Call” the Tactic")
    A --> A2("Ignore the Tactic")
    A --> A3("Discuss What You See and Offer to Help the Other Party Shift to More Honest Behaviors")
    A --> A4("Respond in Kind")

    classDef UmngNode fill:#fdc600,stroke:#043074,stroke-width:2px;
    class A,A1,A2,A3,A4 UmngNode;
```

##

-   **My personal opinion**:

    -   Read the article [@adler_negotiating_2007]:

        -   Before the Bargaining Begins

        -   During the Bargaining Process

    -   I think the best approach is to follow the recommendations in the section **Before the Bargaining Begins** because they are carried out during the planning stage of a negotiation where there is more time and information that can be collected in order to respond adequately to deceptive tactics.

# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html) and [**$\pi$-thon**](https://www.python.org/about/)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://rstudio.com/products/rstudio/), [**Quarto**](https://quarto.org/) and the authors and maintainers of the packages
[**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

# References {.allowframebreaks}