---
title: Best Practices in Negotiations
author: 
  - Luis Francisco Gómez López
institute: 
  - FAEDIS
date: last-modified
format: 
  beamer:
    colortheme: dolphin
    fonttheme: structurebold
    theme: AnnArbor
    link-citations: true
    colorlinks: true
    toc: true
    slide-level: 2
    section-titles: false
    keep-tex: true
    include-in-header:
      file: 000_tex_files/preamble.tex
fig-cap-location: bottom
tbl-cap-location: top
knitr: 
  opts_chunk: 
    echo: false
    warning: false
    fig-align: center
    out-width: 85%
lang: en
bibliography: conflict_neg.bib
---

```{r}
#| label: libraries

library(tidyverse)
library(knitr)
library(kableExtra)
```

# Please Read Me

## 

-   Check the message **Welcome greeting** published in the News Bulletin Board.

-   Dear student please edit your profile uploading a photo where your face is clearly visible.

-   The purpose of the virtual meetings is to answer questions and not to make a summary of the study material.

-   This presentation is based on [@lewicki_negociacion_2024, Chapter 20]

# Purpose

##

Explore best practices in negotiation and conflict management to achieve successful negotiation.

# Final Exam Three Party Coalition Exercise Modified

##

-   Why the Final Exam is a simulations and with the participation of the entire course?

    -   The negotiation of conflicts is generated between individuals or groups and one way of learning is precisely by negotiating with other people.

    -   It is not effective to learn individually and only theoretically.

    -   It is as if a person learned theoretically to play football and without ever playing in a team. Most likely, that person will not perform well in a real match.

-   Before taking part, students should review the general and specific instructions of the Final Exam that can be checked at:

    -   Tercer corte 40% \> Learning Activities \> Final Exam Three Party Coalition Exercise Modified

##

-   Also check out in the **Links of interest** the videos[^1]

    -   *Three-Party Coalition Exercise* [@program_on_negotiation_negotiation_2014] and [@program_on_negotiation_negotiation_2014-1]

[^1]: The videos are in english and are recordings of the Three-Party Coalition Exercise simulation

##

-   Before the Final Exam begins each student of the group, that has been formed, will be randomly assigned to one and only one role as a negotiator of an organization. The respective roles are:

    -   **Group A**
    -   **Group B**
    -   **Group C**

-   The objective of the negotiation is to obtain the highest number of points and determine how they will be divided. This will be reflected in the grade obtained by each student.

##

-   If an agreement is not reached between the parties of the negotiation, each **Group** obtains **0 points** and the grade for each student will be **20** out of 50:

```{r}
#| label: tbl-no-agreement
#| tbl-cap: Results in case of no agreement

tibble(Group  = c("A", "B", "C"),
       Points = rep(x = 0, times = 3),
       Grade  = rep(x = 20, times = 3)) %>% 
  kable(format = "latex", booktabs = TRUE) %>% 
  kable_styling(font_size = 9, latex_options = "striped") %>%
  row_spec(row = 0, bold = TRUE)
```

##

-   If an agreement is reached, it can be obtained between 2 or 3 **Groups**:

    -   Possible agreements:

        -   **Case 1**: **A** and **B** decide to reach an agreement to work together, they obtain **118 points** and must decide how to distribute these points. However, **C** will be excluded.

        -   **Case 2**: **A** and **C** decide to reach an agreement to work together, they obtain **84 points** and must decide how to distribute these points. However, **B** will be excluded.

        -   **Case 3**: **B** and **C** decide to reach an agreement to work together, they obtain **50 points** and must decide how to distribute these points. However, **A** will be excluded.

        -   **Case 4**: **A**, **B** and **C** decide to reach an agreement to work together, they obtain **121 points** and must decide how to distribute these points. Nobody is excluded.

##

-   Grades

    -   **Case 1**: **A** and **B** work together but **C** is excluded.

        -   **C** obtains a grade of **35** out of 50.
        -   The grade of **A** and **B** will depend on who gets the most of the **118 points**. The **Group** that gets the most points will have a grade of **50** out of **50** and the other **Group** gets a grade of **40** out of **50**. If **A** and **B** divide the points equally, whoever gets the highest grade will be assigned randomly.

    -   **Case 2**: **A** and **C** work together but **B** is excluded.

        -   **B** obtains a grade of **35** out of 50.
        -   The grade of **A** and **C** will depend on who gets the most of the **84 points**. The **Group** that gets the most points will have a grade of **50** out of 50 and the other **Group** gets a grade of **40** out of 50. If **A** and **C** divide the points equally, whoever gets the highest grade will be assigned randomly.

##

-   Grades

    -   **Case 3**: **B** and **C** work together but **A** is excluded.

        -   **A** obtains a grade of **35** out of 50.
        -   The grade of **B** and **C** will depend on who gets the most of the **50 points**. The **Group** that gets the most points will have a grade of **50** out of 50 and the other **Group** gets a grade of **40** out of 50. If **B** and **C** divide the points equally, whoever gets the highest grade will be assigned randomly.

    -   **Case 4**: **A**, **B** and **C** work together so nobody is excluded.

        -   The **Group** that obtains the highest amount of points will have a grade of **50** out of 50, the **Group** that obtains the second highest amount of points obtains a grade of **40** out of 50 and the **Group** that obtains the lowest amount of points obtains a grade of **35** out of 50. In case of a tie between any of the **Groups**, the one who obtains the highest grade, the second highest grade or the lowest grade will be assigned randomly depending on whether there is a tie between 2 or 3 **Groups**.

##

-   Before, during and after the Final Exam remember:

    -   **Before**

        -   To coordinate **with all the participants of the course or those who eventually attend** a single group since the activity will consist of a great negotion role play where each student will adopt a role.

        -   To decide whether to appoint one representative per **Group** and the remaining students act as **constituents** or if all the members of the **Group** participate in the discussion.

##

-   Before, during and after the Final Exam remember:

    -   **During**

        -   You have to make 2 decisions: Who do you want to work with? How will the points be divided?
        -   Your grade depends on the amount of points you obtain and no extra points will be assigned for helping or harming the parties involved in the negotiation.
        -   If an agreement is reached and the same amount of points is obtained as another **Group** then the highest grade will be assigned randomly within the **Groups** that obtained the same amount of points.
        -   It is okay to discuss but you must respect the parameters indicated in the last paragraph of the specific instructions.

    -   **After**

        -   Once the negotiation is over, which should last a maximum of 30 minutes, inform the professor of the final result: Was an agreement reached? What was the agreement?

# Best Practices in Negotiations

##

```{dot}
//| label: fig-best-practices-for-negotiators-1
//| fig-cap: Best Practices for Negotiators [@lewicki_negociacion_2024, pp 557-565]
//| fig-width: 4.5
//| fig-height: 2.5

digraph boxes_and_circles {
      
      graph [rankdir = LR]
      
      node[shape = circle, color = "#2C3E50"]
      
      "Best\nPractices";
      
      node[shape = oval, color = "#E31A1C"]
      
      Planning;
      "Structure of\nthe Negotiation";
      BATNA;
      "Negotiation is\nnot necessarily the\nonly option";
      "Handling the apparently opposite\naspects of negotiation";
      
      node[shape = oval, color = "#18BC9C"]
      
      "Own aspects"
      "Other parties aspects"
      "Distributive";
      "Integrative";
      "Blend of both";
      "Claiming value vs Creating Value";
      "Core principles vs Flexible thinking"
      "Being Open vs Being Closed";
      "Being Trusting vs Being Distrusting";
      
       "Best\nPractices" -> Planning;
       "Best\nPractices" -> "Structure of\nthe Negotiation";
       "Best\nPractices" -> BATNA;
       "Best\nPractices" -> "Negotiation is\nnot necessarily the\nonly option";
       "Best\nPractices" -> "Handling the apparently opposite\naspects of negotiation";
       
       Planning -> "Own aspects"
       Planning -> "Other parties aspects"
      "Structure of\nthe Negotiation" -> "Distributive";
      "Structure of\nthe Negotiation" -> "Integrative";
      "Structure of\nthe Negotiation" -> "Blend of both";
      "Handling the apparently opposite\naspects of negotiation" -> "Claiming value vs Creating Value";
      "Handling the apparently opposite\naspects of negotiation" -> "Core principles vs Flexible thinking"
      "Handling the apparently opposite\naspects of negotiation" -> "Being Open vs Being Closed";
      "Handling the apparently opposite\naspects of negotiation" -> "Being Trusting vs Being Distrusting";
}
```

##

```{dot}
//| label: fig-best-practices-for-negotiators-2
//| fig-cap: Best Practices for Negotiators [@lewicki_negociacion_2024, pp 557-565]
//| fig-width: 4.5
//| fig-height: 2.5

digraph boxes_and_circles {
      
      graph [rankdir = LR]
      
      node[shape = circle, color = "#2C3E50"]
      
      "Best\nPractices";
      
      node[shape = oval, color = "#E31A1C"]
      
      "Remember intangible\nfactors";
      "Managing Coalitions";
      "Protect your reputation";
      "Build fairness upon principles"
      "Continue to Learn\nfrom Your Experience"
      
      node[shape = oval, color = "#18BC9C"]
      
      "Against You";
      "For You";
      "Unknown";
      "Establish clear principles";
      "Use external benchmarks";
      "Reach consensus about\nstandards of fairness"
      
       "Best\nPractices" -> "Remember intangible\nfactors";
       "Best\nPractices" -> "Managing Coalitions";
       "Best\nPractices" -> "Protect your reputation";
       "Best\nPractices" -> "Build fairness upon principles";
       "Best\nPractices" -> "Continue to Learn\nfrom Your Experience"
       
       "Managing Coalitions" -> "Against You";
       "Managing Coalitions" -> "For You";
       "Managing Coalitions" -> "Unknown"; 
      
      "Build fairness upon principles" -> "Establish clear principles";
      "Build fairness upon principles" -> "Use external benchmarks";
      "Build fairness upon principles" -> "Reach consensus about\nstandards of fairness"
      
}
```

# What next?

##

-   Book

    -   Raiffa, H., & Richardson, J. (2007). Negotiation analysis: The science and art of collaborative decision making. Cambridge (Mass.: Harvard University Press.

-   Online resource

    -   The Program on Negotiation (PON): <https://www.pon.harvard.edu/>

##

-   Postgraduate education:

    -   **UMNG**: <https://www.umng.edu.co/programas/posgrados/maestria-en-relaciones-y-negocios-internacionales>

        -   **Maestría en Relaciones y Negocios Internacionales**: Negociación Internacional

    -   **Uniandes**: <https://administracion.uniandes.edu.co/programas/posgrados/especializacion-negociacion/>

        -   **Especialización en Negociación**: programa de la Facultad de Administración

# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://rstudio.com/products/rstudio/), [**Quarto**](https://quarto.org/) and the authors and maintainers of the packages [**tidyverse**](https://CRAN.R-project.org/package=tidyverse), [**knitr**](https://CRAN.R-project.org/package=knitr), [**kableExtra**](https://CRAN.R-project.org/package=kableExtra), [**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

-   To the [**Linux kernel community**](https://www.kernel.org/category/about.html) for allowing me the possibility to use some [**Linux distributions**](https://static.lwn.net/Distributions/) as my main [**OS**](https://en.wikipedia.org/wiki/Operating_system) without paying for a license

# References
